---
title: "20250320_DJ11_Julian_24108_Mouse_VEGF"
author: "David John"
date: "2025-31-20"
output: html_document
editor_options: 
  chunk_output_type: console
---

# 0. Load essential libraries

```{r}

setwd(dirname(rstudioapi::getSourceEditorContext()$path))
cfg<-config::get()
getwd()
Sys.getenv("")

library(reticulate)
use_python("/usr/bin/python3.12")
#load libraries
library(tidyverse)
library(ggplot2)
library(Seurat)
library(openxlsx)
library(DOtools)
source(file.path(cfg$mount_point,"ZMM_shared/Bioinformatic/scRNA-SEQ-ZMM/Mariano_functions.R"))
source(file.path(cfg$mount_point,"ZMM_shared/Bioinformatic/scRNA-SEQ-ZMM/Import10X-HelperFunctions_SeuratV5.R"))
outputFolder <- file.path(cfg$mount_point,"Sequencing_Storage/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT_Elvin/")
sink(file = file.path(outputFolder,"/analysis/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT.log"), append = TRUE, split = TRUE)
```

# 1. Preprocessing data 
## 1.1 Prepare Paths for import count matrixes from cellbender outputs

```{r}

Sample_Paths <- c(paste0(outputFolder,"raw-data/cellbender/Tcf15mutant1/"),
                  paste0(outputFolder,"raw-data/cellbender/Tcf15mutant2/"),
                  paste0(outputFolder,"raw-data/cellbender/Tcf15mutant3"),
                  paste0(outputFolder,"raw-data/cellbender/Wildtype1/"),
                  paste0(outputFolder,"raw-data/cellbender/Wildtype2/"),
                  paste0(outputFolder,"raw-data/cellbender/Wildtype3/")
                  )




Sample_Names <- c("Tcf15mutant-1","Tcf15mutant-2","Tcf15mutant-3","Wildtype-1", "Wildtype-2", "Wildtype-3")
```

## 1.2 Run Importer function

```{r}
#import data, create SeuratObjects
resultList<-list()
SeuratObjectList <- list()
for (i in 1:length(Sample_Paths)) {
  resultList<-Importer(pathway = Sample_Paths[i],
                       id = Sample_Names[i],
                       TenX = FALSE,
                       CellBender = TRUE,
                       FilterCells = TRUE,
                       FilterByAbsoluteValues = TRUE,
                       DeleteDoublets = TRUE,
                       minFeatures=300, 
                       maxFeatures=6000,
                       minCounts=500,
                       maxCounts=20000,
                       performScaling = TRUE,
                       maxMito=0.05) # warnings result from not specified layer in Vlnplot function, it still works correctly but annoying
  print(resultList[[2]])
  SeuratObjectList[[i]]<-resultList[[1]]
}

```

## 1.3 Add disease information to the objects before integrating

```{r}
SeuratObject.anchors <- FindIntegrationAnchors(object.list = SeuratObjectList, dims = 1:20)
SeuratObject.combined <- IntegrateData(anchorset = SeuratObject.anchors, dims = 1:20)
```

## 1.4 Integration with CCA method

```{r}
merged_obj <- Reduce(function(x, y) merge(x, y), SeuratObjectList) # get rid of the error prone approach

merged_obj <- NormalizeData(merged_obj)
merged_obj <- FindVariableFeatures(merged_obj)
merged_obj <- ScaleData(merged_obj)
merged_obj <- RunPCA(merged_obj)

SeuratObject.combined <- IntegrateLayers(object = merged_obj, method = CCAIntegration, orig.reduction = "pca", new.reduction = "integrated.cca", verbose = T, )

#re-join layers after integration
SeuratObject.combined <- JoinLayers(SeuratObject.combined) # the individual scale layers remain after joining layers, remove them they are just unnecessary memory waste 
layers_to_remove <- grep("^scale\\.data\\.", Layers(SeuratObject.combined), value = TRUE)
SeuratObject.combined@assays$RNA@layers[layers_to_remove] <- NULL

saveRDS(SeuratObject.combined, file = "/mnt/david/Sequencing_Storage/20250429_DJ016_snRNASEQ_25034_Mouse_Elvin_wholeHeart_Apex_TCF15flox_vs_WT/analysis/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT.rds")
```

# 2. Calculate neighbors, clusters and umap representation

```{r}
SeuratObject.combined <- FindNeighbors(object = SeuratObject.combined, reduction = "integrated.cca", dims = 1:10)
SeuratObject.combined <- FindClusters(SeuratObject.combined, resolution = 0.6, algorithm=1, random.seed = 1234)
SeuratObject.combined <- RunUMAP(object = SeuratObject.combined,
                     reduction = "integrated.cca",
                     dims = 1:10,
                     spread = 1,
                     min.dist = 0.3)



p <- DimPlot(SeuratObject.combined, group.by = "seurat_clusters", label = T, raster = F) +
    labs(x="UMAP1",y = "UMAP2")+
    theme(plot.title = element_blank(),
          # text = element_text(face = "bold",size = 20),
          axis.title.x = element_text(size = 14, family="Helvetica"),
          axis.title.y = element_text(size = 14, family="Helvetica"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          legend.text = element_text(face = "bold"))

ggsave(paste0(outputFolder, "analysis/plots/global/UMAPs/UMAP_seurat_cluster.pdf"), width = 8,height = 8, plot = p)

SeuratObject.combined <- DO.CellTypist(SeuratObject.combined, runCelltypistUpdate=F)
SeuratObject.combined$annotation <- SeuratObject.combined$predicted_labels_celltypist

p <- DimPlot(SeuratObject.combined, group.by = "predicted_labels_celltypist", label = T, cols = umap_colors, raster = F) +
    labs(x="UMAP1",y = "UMAP2")+
    theme(plot.title = element_blank(),
          # text = element_text(face = "bold",size = 20),
          axis.title.x = element_text(size = 14, family="Helvetica"),
          axis.title.y = element_text(size = 14, family="Helvetica"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          legend.text = element_text(face = "bold"))

ggsave(paste0(outputFolder, "analysis/plots/global/UMAPs/UMAP_seurat_celltypes.pdf"), width = 8,height = 8, plot = p)

saveRDS(SeuratObject.combined, file = "/mnt/david/Sequencing_Storage/20250429_DJ016_snRNASEQ_25034_Mouse_Elvin_wholeHeart_Apex_TCF15flox_vs_WT/analysis/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT.rds")

SeuratObject.combined <- readRDS(file.path(cfg$mount_point,"Sequencing_Storage/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT_Elvin/analysis/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT.rds"))

```

## 2.1 Refine Annotation

```{r}
#Reclustering using FindSubCLuster in a loop
SeuratObject.combined$seurat_Recluster <- as.vector(SeuratObject.combined$seurat_clusters)
for (cluster in unique(SeuratObject.combined$seurat_clusters)) {
  print(cluster)
  SeuratObject.combined <- FindSubCluster(SeuratObject.combined,
                            cluster = cluster,
                            graph.name = "RNA_snn",
                            algorithm = 4,
                            resolution = 0.3)  
  
  cluster_cells <- rownames(SeuratObject.combined@meta.data)[SeuratObject.combined$seurat_clusters == cluster]
  SeuratObject.combined$seurat_Recluster[cluster_cells] <- SeuratObject.combined$sub.cluster[cluster_cells]
}
warSeuratObject.combined$sub.cluster <- NULL

#One Liner
# SeuratObject <- Seu_miRNA
# SeuratObject <- DO.FullRecluster(SeuratObject = SeuratObject,
#                  res = 0.5,
#                  algorithm = 4,
#                  graph.name = "RNA_snn")

p <- DimPlot(SeuratObject.combined, group.by = "seurat_Recluster", label = T, cols = rep(umap_colors,5), raster = F) +
    labs(x="UMAP1",y = "UMAP2")+
    theme(plot.title = element_blank(),
          # text = element_text(face = "bold",size = 20),
          axis.title.x = element_text(size = 14, family="Helvetica"),
          axis.title.y = element_text(size = 14, family="Helvetica"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          legend.text = element_text(face = "bold"))

ggsave(paste0(outputFolder, "analysis/plots/global/UMAPs/UMAP_seurat_Recluster.pdf"), width = 8,height = 8, plot = p)


#re-run annotation
SeuratObject.combined <- DO.CellTypist(SeuratObject.combined, over_clustering = "seurat_Recluster", runCelltypistUpdate=F)
SeuratObject.combined$annotation_reclustered <- SeuratObject.combined$predicted_labels_celltypist
p <- DimPlot(SeuratObject.combined, group.by = "annotation_reclustered", label = T, cols = umap_colors, raster = F) +
    labs(x="UMAP1",y = "UMAP2")+
    theme(plot.title = element_blank(),
          # text = element_text(face = "bold",size = 20),
          axis.title.x = element_text(size = 14, family="Helvetica"),
          axis.title.y = element_text(size = 14, family="Helvetica"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          legend.text = element_text(face = "bold"))

ggsave(paste0(outputFolder, "analysis/plots/global/UMAPs/UMAP_predicted_cell_annotation_reclustered.pdf"), width = 8,height = 8, plot = p)

SeuratObject.combined$annotation_refined <- SeuratObject.combined$annotation_reclustered
table(SeuratObject.combined$annotation_reclustered)

SeuratObject.combined$annotation_refined <- plyr::revalue(SeuratObject.combined$annotation_reclustered,
                                              c(
                                                `B` = "B_cells",
                                                `CD16+Mo` = "Ccr2+MP",
                                                `CD4+T_naive` = "T_cells", # Cd4+ and Cd8+
                                                `EC1_cap` = "Cap_EC",
                                                `EC5_art` = "Art_EC",
                                                `EC7_endocardial` = "End_EC",
                                                `EC8_ln` = "Lym_EC",
                                                `FB1` = "FB",
                                                `FB2` = "FB",
                                                `FB5` = "FB",
                                                `FB4_activated` = "Act_FB",
                                                `LYVE1+IGF1+MP` = "LYVE1+MP",
                                                `Meso` = "Epic",
                                                `NC1_glial` = "Neuro",
                                                `NC2_glial_NGF+` = "PC",
                                                `PC3_str` = "PC",
                                                `SMC1_basic` = "SMC",
                                                `vCM1` = "CM1",
                                                `vCM4` = "CM2"))

table(SeuratObject.combined$annotation_refined)

p <- DimPlot(SeuratObject.combined, group.by = "annotation_refined", label = T, cols = umap_colors, raster = F) +
    labs(x="UMAP1",y = "UMAP2")+
    theme(plot.title = element_blank(),
          # text = element_text(face = "bold",size = 20),
          axis.title.x = element_text(size = 14, family="Helvetica"),
          axis.title.y = element_text(size = 14, family="Helvetica"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          legend.text = element_text(face = "bold"))

ggsave(paste0(outputFolder, "analysis/plots/global/UMAPs/UMAP_Refined_cell_annotation.pdf"), width = 8,height = 8, plot = p)

 saveRDS(SeuratObject.combined, file = "/mnt/david/Sequencing_Storage/20250429_DJ016_snRNASEQ_25034_Mouse_Elvin_wholeHeart_Apex_TCF15flox_vs_WT/analysis/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT.rds")

SeuratObject.combined <- readRDS(file = "/Volumes/Sequencing_Storage/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT_Elvin/analysis/20250429_DJ016_snRNASEQ_25034_Mouse_wholeHeart_Apex_TCF15flox_vs_WT.rds")


```

## 2.2 UMAP per Condition

```{r}
p <- DimPlot(SeuratObject.combined, group.by = "annotation_refined", label = T, cols = umap_colors, raster = F, split.by = "condition") +
    labs(x="UMAP1",y = "UMAP2")+
    theme(plot.title = element_blank(),
          # text = element_text(face = "bold",size = 20),
          axis.title.x = element_text(size = 14, family="Helvetica"),
          axis.title.y = element_text(size = 14, family="Helvetica"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          legend.position = "none",
          legend.text = element_text(face = "bold"))
ggsave(paste0(outputFolder, "analysis/plots/global/UMAPs/UMAP_Refined_cell_annotation_byCondition.pdf"), width = 8,height = 8, plot = p)
```

##2.3 Barplot of nuclei per sample
```{r}
# orig.ident is Seurat's default sample identifier
# 1. Nuclei pro Sample zählen
cell_counts <- SeuratObject.combined@meta.data %>%
  group_by(orig.ident, condition) %>%
  summarise(Nuclei = n(), .groups = "drop")

# 2. Statistiken pro Condition berechnen
stats <- cell_counts %>%
  group_by(condition) %>%
  summarise(
    mean_nuclei = mean(Nuclei),
    sd_nuclei   = sd(Nuclei),
    .groups = "drop"
  )

#output raw values
openxlsx::write.xlsx(cell_counts, file = file.path(outputFolder, "analysis/excels/qc_nuclei_per_condition.xlsx"))
openxlsx::write.xlsx(stats, file = file.path(outputFolder, "analysis/excels/qc_nuclei_per_condition_average.xlsx"))

# Set the desired order of conditions
cell_counts$condition <- factor(cell_counts$condition, levels = c("Wildtype", "Tcf15mutant"))
stats$condition       <- factor(stats$condition,       levels = c("Wildtype", "Tcf15mutant"))

# 3. Plot
ggplot() +
  # Balken (Mittelwert)
  geom_bar(data = stats,
           aes(x = condition, y = mean_nuclei, fill = condition),
           stat = "identity", width = 0.6) +
  # Error Bars (SD)
  geom_errorbar(data = stats,
                aes(x = condition,
                    ymin = mean_nuclei - sd_nuclei,
                    ymax = mean_nuclei + sd_nuclei),
                width = 0.2, linewidth = 0.8) +
  # Einzelpunkte (jittered)
  geom_jitter(data = cell_counts,
              aes(x = condition, y = Nuclei),
              width = 0.1, size = 2.5, shape = 21,
              fill = "black", color = "black", stroke = 0.8) +
  # Farben: beige & blau
  scale_fill_manual(values = c("#F5F0DC", "#4472C4")) +
  theme_classic() +
  labs(title = "Nuclei per Condition",
       x = "Condition",
       y = "Number of Nuclei") +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1),
        legend.position = "none", plot.title = element_text(hjust = 0.5))

ggsave(filename = file.path(outputFolder, "analysis/plots/barplots/qc_nuclei_per_condition_barplot.svg"))
```


# 3. Define Markers for each cluster

```{r}
#Calculate Top expressed Markers for annotation
Idents(SeuratObject.combined)<-"annotation_refined"
SeuratObject.combined<-JoinLayers(SeuratObject.combined)
annotation_Markers<- FindAllMarkers(object = SeuratObject.combined,
                                    assay = "RNA",
                                    group.by = "annotation_refined",
                                    min.pct = 0.25,
                                    logfc.threshold = 0.25)

top_markers <- annotation_Markers %>%
  dplyr::group_by(cluster) %>%
  top_n(n = 5, wt = avg_log2FC)

#kick MT-genes
MT.genes <- top_markers[grep(pattern = "MT-", top_markers$gene),]
top_markers <- top_markers[!(top_markers$gene %in% MT.genes$gene),]

DO.Dotplot(object = SeuratObject.combined,
           Feature = top_markers,
           group.by.x = "annotation_refined",
           plot.margin = c(1,1,1,1),
           annotation_x = T, # only useable if Feature is a dataframe with cluster information, e.g. FindAllMakers output (you need a column with cluster)
           point_stroke=0.1,
           annotation_x_rev = T,
           textSize = 14,
           hjust = 0.5,
           vjust = 0,
           textRot = 0,
           segWidth = 0.3,
           lwd = 3
        )

ggsave(paste0(outputFolder,"analysis/plots/dotplots/TopExpressed_genes_per_celltype.svg"), width = 19, height = 10)
ggsave(paste0(outputFolder,"analysis/plots/dotplots/TopExpressed_genes_per_celltype.pdf"), width = 19, height = 10)

```

# 4. DEG analysis per cell type

```{r}
DEG_stats_collector <- list()
table(SeuratObject.combined$condition)

#Define Control and Disease condition
ctrl.condition<-"Wildtype"
dis.condition<-c("Tcf15mutant")
for (dis in dis.condition) {
  print(dis)
 for (celltype in unique(SeuratObject.combined$annotation_refined)) {
   print(celltype)
   Seu_celltype <- subset(SeuratObject.combined, subset = annotation_refined == celltype) # for DEG analysis the normal subset function should still work
  
  DEG_stats <- FindMarkers(object = Seu_celltype,
                           ident.1 = dis,
                           ident.2 = ctrl.condition,
                           min.pct = 0.01,
                           logfc.threshold = 0.1,
                           assay = "RNA",
                           only.pos = F,
                           test.use = "wilcox",
                           group.by = "condition")
  DEG_stats <- rownames_to_column(DEG_stats)
  DEG_stats[["condition"]] <- dis
  DEG_stats[["celltype"]] <- celltype
  
  DEG_stats_collector <- rbind(DEG_stats_collector, DEG_stats)     
  
 }
}

write.xlsx(DEG_stats_collector, file = paste0(outputFolder, "analysis/excels/DEG_analysis_nothresh_TCF15_vs_WT.xlsx"))
```

## 4.1 Dotplots per Celltype

```{r}
top_markers <- DEG_stats_collector %>%
  dplyr::group_by(celltype) %>%
  top_n(n = 5, wt = abs(avg_log2FC))

top_markers <- DEG_stats_collector %>%
  arrange(desc(abs(avg_log2FC))) %>%              
  distinct(rowname, .keep_all = TRUE) %>%       
  group_by(celltype) %>%
  slice_head(n = 5) 

top_markers %>% rename(rowname="gene", celltype="cluster") -> top_markers



#kick MT-genes
MT.genes <- top_markers[grep(pattern = "MT-", top_markers$gene),]
top_markers <- top_markers[!(top_markers$gene %in% MT.genes$gene),]

library(openxlsx)
write.xlsx(top_markers, file = paste0(outputFolder, "analysis/excels/DEG_analysis_top20_DEGs_perCelltype_by_log2FC_TCF15_vs_WT.xlsx"))


top_markers_uniq<- top_markers %>% distinct(gene, .keep_all = FALSE)
DO.Dotplot(SeuratObject.combined, Feature = top_markers, group.by.x = "condition", annotation_x = TRUE, segWidth=0.3, hjust=0.5, vjust=0, lwd=3)
ggsave(paste0(outputFolder,"analysis/plots/global/dotplots/Top5Expressed_genes_TCF15_vs_WT.pdf"), width = 19, height = 10)

celltypes<-unique(SeuratObject.combined$annotation_refined)
for (celltype in unique(SeuratObject.combined$annotation_refined)) {
   print(celltype)

  df.tmp<-top_markers[top_markers$cluster==celltype,]  
   
}



DO.Dotplot(SeuratObject.combined, Feature = top_markers, group.by.x = "condition")
DoHeatmap(SeuratObject.combined, features = top_markers$gene, group.by = "condition")
```

## 4.2 Plot TCF15 Expression

```{r}
library(DOtools)
table(SeuratObject.combined$condition)
p<-DO.BarplotWilcox(Seu_object = SeuratObject.combined, Feature = "Tcf15", ctrl.condition = "Wildtype")
ggsave(filename = paste0(outputFolder, "analysis/plots/global/vlnplots/Barplot_Tcf15_global.pdf"), width = 5, height = 5, plot = p, device = "pdf")

```

```{r}
FeaturePlot(SeuratObject.combined, features = "Tcf15", label = T, label.size = 4, cols = c("grey", "red"), pt.size = 3, min.cutoff = 1, max.cutoff = 2)
ggsave(filename = paste0(outputFolder, "analysis/plots/global/UMAPs/FeaturePlot_Tcf15_global.pdf"), width = 5, height = 5,  device = "pdf")
```

```{r}

DO.UMAP(sce_object = SeuratObject.combined, features = "Tcf15", group.by = "annotation", label = TRUE, DensityPlot = TRUE, FeaturePlot = FALSE)
ggsave(filename = paste0(outputFolder, "analysis/plots/global/UMAPs/DensityPlot_Tcf15_global.pdf"), width = 5, height = 5,  device = "pdf")
```


#### 4.3 Claculate Mesenchmyal score

```{r}
mesenchymal_genes_for_score<-list(mesenchymal=c("Mmp14","Fn1","Postn","Eln","Col1a1","Col1a2","Mgp","Col3a1","Bgn","Dcn","Serpine1","Mmp9","Snai1","Egfl7"))
SeuratObject.combined<-AddModuleScore(object = SeuratObject.combined, features = mesenchymal_genes_for_score, name = "mesenchymal_score")

FeaturePlot(SeuratObject.combined, features = "mesenchymal_score1", label = T, label.size = 4, split.by = "condition")
ggsave(filename = paste0(outputFolder, "analysis/plots/global/UMAPs/FeaturePlot_MesenchymalScore_global.pdf"), width = 8, height = 5,  device = "pdf")
DO.BarplotWilcox(SeuratObject.combined, Feature = "mesenchymal_score1", ctrl.condition = "Wildtype")
ggsave(filename = paste0(outputFolder, "analysis/plots/global/vlnplots/Barplot_MesenchymalScore_global_wilcox.pdf"), width = 5, height = 5,  device = "pdf")
```

# 5  Cell compositions

## 5.1 stacked Barplot

```{r}
# get information of cell type number of object
barplot.DF <- SeuratObject.combined@meta.data %>%
  group_by(cell.type = annotation_refined, condition) %>%
  summarise(value = n(), .groups = "drop")

barplot.DF$condition <- factor(barplot.DF$condition)

# Convert to a contingency table
contingency_table <- pivot_wider(
  data = barplot.DF, 
  names_from = condition, 
  values_from = value,
  values_fill = 0
) %>% 
  column_to_rownames("cell.type")

write.xlsx(contingency_table, paste0(outputFolder,"analysis/excels/Cell_compositions_numbers_perCondition.xlsx"), rowNames=T)

chisq_result <- chisq.test(contingency_table)
p_value <- signif(chisq_result$p.value, digits = 3) 
p_value <- signif(ifelse(p_value == 0, .Machine$double.xmin, p_value),digits = 3) 

ggplot(barplot.DF, aes(fill=cell.type, y=value,x=condition))+
  geom_bar(position = "fill", stat= "identity")+
  scale_fill_manual(values = umap_colors
                      , name = "Cell type")+
  labs(title = "", y = "Cell composition (in %)") +
  theme_classic() +
  theme(axis.text.x = element_text(color = "black",angle = 0, size = 16),
        axis.text.y = element_text(color = "black", size = 16),
        axis.title.x = element_blank(),
        axis.title = element_text(size = 16, color = "black"),
        plot.title = element_text(size = 16, hjust = 0.5),
        axis.line = element_line(color = "black"),
        strip.text.x = element_text(size = 16, color = "black"),
        legend.position = "right") 
  annotate("text", x = 1.5, y = 1.05, label = paste("p =", p_value), size = 6, color = "black")

ggsave(paste0(outputFolder, "analysis/plots/barplots/Stacked_bar_plot_cell_composition.pdf"), width = 5, height = 5)
```

## 5.2 Barplot per sample

```{r}
p<-BarplotPercentages(SeuratObject = SeuratObject.combined, split.by = "orig.ident", group.by = "annotation_refined", color.by = "condition")
ggsave(filename = paste0(outputFolder, "analysis/plots/barplots/Barplot_cell_composition_per_celltype_sample.pdf"), plot =p, device = "pdf",  width = 10, height = 10)
```

##5.1 stacked Barplot

```{r}
# get information of cell type number of object
barplot.DF <- SeuratObject.combined@meta.data %>%
  group_by(cell.type = annotation_refined, orig.ident) %>%
  summarise(value = n(), .groups = "drop")

barplot.DF$orig.ident <- factor(barplot.DF$orig.ident)

# Convert to a contingency table
contingency_table <- pivot_wider(
  data = barplot.DF, 
  names_from = orig.ident, 
  values_from = value,
  values_fill = 0
) %>% 
  column_to_rownames("cell.type")

write.xlsx(contingency_table, paste0(outputFolder,"analysis/excels/Cell_compositions_numbers_by_orig.ident.xlsx"), rowNames=T)

chisq_result <- chisq.test(contingency_table)
p_value <- signif(chisq_result$p.value, digits = 3) 
p_value <- signif(ifelse(p_value == 0, .Machine$double.xmin, p_value),digits = 3) 

ggplot(barplot.DF, aes(fill=cell.type, y=value,x=orig.ident))+
  geom_bar(position = "fill", stat= "identity")+
  scale_fill_manual(values = umap_colors
                      , name = "Cell type")+
  labs(title = "", y = "Cell composition (in %)") +
  theme_classic() +
  theme(axis.text.x = element_text(color = "black",angle = 45, size = 16, hjust = 0.9, vjust = 0.9),
        axis.text.y = element_text(color = "black", size = 16),
        axis.title.x = element_blank(),
        axis.title = element_text(size = 16, color = "black"),
        plot.title = element_text(size = 16, hjust = 0.5),
        axis.line = element_line(color = "black"),
        strip.text.x = element_text(size = 16, color = "black"),
        legend.position = "right") 
  #annotate("text", x = 1.5, y = 1.05, label = paste("p =", p_value), size = 6, color = "black")

ggsave(paste0(outputFolder, "analysis/plots/barplots/Stacked_bar_plot_cell_composition_orig_ident.pdf"), width = 5, height = 5)
```

```{r}
p<-DO.Vln.Plot.wilcox(SeuratObject.combined, SeuV5 = T, Feature = "Tcf15", ctrl.condition = "Wildtype", group.by = "condition")
ggsave(filename = paste0(outputFolder, "analysis/plots/vlnplotsVLNPlot_Tcf15_global.pdf"), width = 5, height = 5, plot = p, device = "pdf")

```

# 6. Celltype specific Analyis 
## 6.1vCM1 (renamed to CM1)

```{r}
Seu.CM<-subset(SeuratObject.combined, subset = annotation_refined == "CM1")
```

```{r}
DO.Vln.Plot.wilcox(Seu.CM, SeuV5 = T, Feature = "TCF15", ctrl.condition = "Wildtype", group.by = "condition")
ggsave(paste0(outputFolder, "analysis/plots/VLNPlot_TCF15_CM1.pdf"), width = 5, height = 5)

```

## 6.2 Endothelial cells 
### 6.2.1 All Ecs

```{r}
table(SeuratObject.combined$annotation_refined)
Seu.EC_all<-subset(SeuratObject.combined, subset = annotation_refined %in% c("Cap_EC", "Art_EC", "EC6_ven", "End_EC"))
DO.Dotplot(Seu_object = Seu.EC_all,
           Feature = c("Serpine1","Egfl7","Pecam1", "Twist1", "Snai1"),
           group.by.x = "condition",
           plot.margin = c(1,1,1,1),
           annotation_x = F, # only useable if Feature is a dataframe with cluster information, e.g. FindAllMakers output (you need a column with cluster)
           point_stroke=0.1,
           annotation_x_rev = F,
           textSize = 14,
           hjust = 0.5,
           vjust = 0,
           textRot = 0,
           segWidth = 0.3,
           lwd = 3
        )

DO.Barplot(Seu.EC_all, Feature = "Meox2", ctrl.condition = "Wildtype", group.by = )

.logger<-DOtools:::.logger
genes<-c("Serpine1","Egfl7","Pecam1", "Twist1", "Snai1")
for (gene in genes) {
  DO.Vln.Plot.wilcox(Seu.EC_all, SeuV5 = T, Feature = gene, ctrl.condition = "Wildtype", group.by = "condition")
  
  filename<-paste0()
  ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/VLNPlot_",gene,"_EC_all.pdf"), width = 5, height = 5)
}

for (gene in genes) {
  DO.Mean.SEM.Graphs.cluster.t(Seu.EC_all, Feature = gene, ctrl.condition = "Wildtype", group.by = "condition")
  filename<-paste0()
  ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/Mean.SEM.cluster.tTest_",gene,"_EC_all.pdf"), width = 5, height = 5)
  
  DO.Mean.SEM.Graphs.wilcox(Seu.EC_all, Feature = gene, ctrl.condition = "Wildtype", group.by = "condition")
  filename<-paste0()
  ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/Mean.SEM.wilcoxon_",gene,"_EC_all.pdf"), width = 5, height = 5)
}

```


### 6.2.2 Plot transporter genes in EC subtypes ()
```{r}
DO.Dotplot(object = Seu.EC_all,
           Feature = c("Cd36","Slc2a1","Cav1", "Insr"),
           group.by.x = "condition",
           group.by.y = "annotation_refined",
           plot.margin = c(1,1,1,1),
           annotation_x = F, # only useable if Feature is a dataframe with cluster information, e.g. FindAllMakers output (you need a column with cluster)
           point_stroke=0.1,
           annotation_x_rev = F,
           textSize = 14,
           hjust = 0.5,
           vjust = 0,
           textRot = 0,
           segWidth = 0.3,
           lwd = 3, across.group.by.x = "condition"
          
           
        )


genes<- c("Cd36","Slc2a1","Cav1", "Insr")
for (gene in genes){
  p<- DO.BarplotClustert(Seu.EC_all, Feature = gene, ctrl.condition = "Wildtype")
  p<- ggpubr::annotate_figure(p, bottom = "p-value= t-test clustered over samples")

  ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/BarPlot_",gene,"clustered_t_test_EC_all.pdf"), width = 5, height = 5)
  
}

for (celltype in unique(Seu.EC_all$annotation_refined)) {
  for (gene in genes){
    sub <- subset(Seu.EC_all, subset = annotation_refined == celltype)
    p<- DO.BarplotClustert(sub, Feature = gene, ctrl.condition = "Wildtype")
    p<- ggpubr::annotate_figure(p, top = celltype, bottom = "p-value= t-test clustered over samples")
    ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/BarPlot_",celltype,"_",gene,"_clustered_t_test_EC_all.pdf"), width = 5, height = 5)
  }
}
p<- DO.BarplotClustert(Seu.EC_all, Feature = "Cav1", ctrl.condition = "Art_EC", group.by = "annotation_refined",  )

```


```{r}
genes<- c("Meox2")

for (celltype in unique(Seu.EC_all$annotation_refined)) {
  for (gene in genes){
    sub <- subset(Seu.EC_all, subset = annotation_refined == celltype)
    p<- DO.BarplotClustert(sub, Feature = gene, ctrl.condition = "Wildtype")
    p<- ggpubr::annotate_figure(p, top = celltype, bottom = "p-value= t-test clustered over samples")
    ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/BarPlot_",celltype,"_",gene,"_clustered_t_test_EC_all.pdf"), width = 5, height = 5)
  }
}

```


#### 6.2.1.1 Plot Mesenchymal genes and Score

```{r}
mesenchymal_genes1<-c("Fn1","Eln","Mgp","Bgn","Mmp14", "Dcn","Col3a1")
mesenchymal_genes2<-c("Serpine1","Mmp9","Snai1", "Postn","Col1a1","Col1a2")

vascular_genes<-c("Pecam1","Kdr","Ptprb","Spry1","Foxo1","S1pr1","Efna1","Gja4","Cxcr4","Ace","Tgfbr2","Irp5")

#rename Wildtype to Ctrl, so that it apperas first in the dotplot
Seu.EC_all$condition2<- ifelse(Seu.EC_all$condition=="Wildtype", yes = "Control", no = "Tcf15mutat")


p<- DO.Dotplot(Seu_object = Seu.EC_all,
           Feature = mesenchymal_genes1,
           group.by.x = "condition2",
           group.by.y = "annotation_refined",
           plot.margin = c(1,1,1,1),
           annotation_x = F, # only useable if Feature is a dataframe with cluster information, e.g. FindAllMakers output (you need a column with cluster)
           point_stroke=0.1,
           annotation_x_rev = F,
           textSize = 14,
           hjust = 0.5,
           vjust = 0,
           textRot = 0,
           segWidth = 0.3,
           lwd = 3, scale_gene = TRUE,
        ) + ggtitle("Mesenchymal Genes in ECs")

ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/dotplot_mesenchymalmarkers_ec_all_1.pdf"), width = 10, height = 4)

p<- DO.Dotplot(Seu_object = Seu.EC_all,
           Feature = mesenchymal_genes2,
           group.by.x = "annotation_refined",
           group.by.y = "condition2",
           plot.margin = c(1,1,1,1),
           annotation_x = F, # only useable if Feature is a dataframe with cluster information, e.g. FindAllMakers output (you need a column with cluster)
           point_stroke=0.1,
           annotation_x_rev = F,
           textSize = 14,
           hjust = 0.5,
           vjust = 0,
           textRot = 0,
           segWidth = 0.3,
           lwd = 3, scale_gene = TRUE,
        ) + ggtitle("Mesenchymal Genes in ECs")

ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/dotplot_mesenchymalmarkers_ec_all_2.pdf"), width = 10, height = 4)

 p<- DO.Dotplot(Seu_object = Seu.EC_all,
           Feature = vascular_genes,
           group.by.x = "annotation_refined",
           group.by.y = "condition",
           plot.margin = c(1,1,1,1),
           annotation_x = F, # only useable if Feature is a dataframe with cluster information, e.g. FindAllMakers output (you need a column with cluster)
           point_stroke=0.1,
           annotation_x_rev = F,
           textSize = 14,
           hjust = 0.5,
           vjust = 0,
           textRot = 0,
           segWidth = 0.3,
           lwd = 3, 
        ) + ggtitle("Vascular Genes in ECs")

ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/dotplot_vascularmarkers_ec_all.pdf"), width = 15, height = 10)
```

#### 6.2.1.2  Plot EC subtypes separately (Fig 3e (Final))
```{r}
library(ggpubr)
SeuratObject.combined$condition2<- ifelse(SeuratObject.combined$condition=="Wildtype", yes = "Ctrl", no = "Tcf15mutat")
mesenchymal_genes1<-c("Fn1","Eln","Mgp","Bgn","Mmp14","Dcn","Col3a1")
mesenchymal_genes2<-c("Serpine1","Mmp9","Snai1","Postn","Col1a1","Col1a2")

for (celltype in c("Cap_EC", "Art_EC", "EC6_ven", "End_EC")) {
  Seu.EC_temp <-subset(SeuratObject.combined, subset= annotation_refined == celltype)


  #p1<- DO.Dotplot(Seu_object = Seu.EC_temp, Feature = mesenchymal_genes1, group.by.x = "condition2") + ggtitle(paste("Mesenchymal Genes in ", celltype)) + coord_flip()
   # ggsave(paste0(outputFolder, "analysis/plots/for-paper/Fig3e/dotplot_mesenchymal_genes1_", celltype, "_.pdf"), width = 5, height = 7)

  
  p2<- DO.Dotplot(Seu_object = Seu.EC_temp, Feature = mesenchymal_genes2, group.by.x = "condition2") + coord_flip()
    ggsave(paste0(outputFolder, "analysis/plots/for-paper/Fig3e/dotplot_mesenchymal_genes2_", celltype, "_.pdf"), width = 5, height = 7)

} 
```


```{r}
library(ggpubr)
SeuratObject.combined$condition2<- ifelse(SeuratObject.combined$condition=="Wildtype", yes = "Control", no = "Tcf15mutat")
mesenchymal_genes1<-c("Fn1","Eln","Mgp","Bgn","Dcn","Col3a1")
mesenchymal_genes2<-c("Serpine1","Mmp9","Snai1","Postn","Col1a1","Col1a2")




SeuratObject.Ecs<-DO.Subset(SeuratObject.combined, subset = annotation_refined %in% c("Cap_EC", "Art_EC", "EC6_ven", "End_EC"))



DO.Heatmap(sce_object = SeuratObject.Ecs, features = c(mesenchymal_genes1,mesenchymal_genes2), group_by = "condition", xticks_rotation = 45, title = "EC_all", path = paste0(outputFolder, "analysis/plots/celltype/EC_all"), filename = "heatmap_mesenchymalmarkers_ec_all.pdf", swap_axes = FALSE) 


DO.Dotplot(object = SeuratObject.Ecs, Feature = mesenchymal_genes1, group.by.x = "annotation_refined", group.by.y = "condition")

for (celltype in c("Cap_EC", "Art_EC", "EC6_ven", "End_EC")) {
  Seu.EC_temp <-subset(SeuratObject.combined, subset= annotation_refined == celltype)

  DO.Heatmap(sce_object = Seu.EC_temp, features = c(mesenchymal_genes1,mesenchymal_genes2), group_by = "condition", xticks_rotation = 45, title = celltype, path = paste0(outputFolder, "analysis/plots/celltype/"), filename = paste0("heatmap_mesenchymalmarkers_",celltype,".pdf"), swap_axes = FALSE) 

  #p1<- DO.Dotplot(Seu_object = Seu.EC_temp, Feature = mesenchymal_genes1, group.by.x = "condition2") + ggtitle(paste("Mesenchymal Genes in ", celltype)) + coord_flip()
   # ggsave(paste0(outputFolder, "analysis/plots/for-paper/Fig3e/dotplot_mesenchymal_genes1_", celltype, "_.pdf"), width = 5, height = 7)

  
  #p2<- DO.Dotplot(Seu_object = Seu.EC_temp, Feature = mesenchymal_genes2, group.by.x = "condition2") + coord_flip()
    #ggsave(paste0(outputFolder, "analysis/plots/for-paper/Fig3e/dotplot_mesenchymal_genes2_", celltype, "_.pdf"), width = 5, height = 7)

} 

sub<-subset(SeuratObject.combined, subset= annotation_refined == "Cap_EC")
DO.Heatmap(sub, features = c("Snai1","Serpine1"), group_by = "condition", xticks_rotation = 45, title = "Cap_EC", path = paste0(outputFolder, "analysis/plots/celltype/"), filename = paste0("heatmap_mesenchymalmarkers_","Cap_EC","_20260617.pdf"), swap_axes = FALSE)
```

#### 6.2.1.2 Claculate Mesenchmyal score

```{r}
mesenchymal_genes_for_score<-list(mesenchymal=c("Mmp14","Fn1","Postn","Eln","Col1a1","Col1a2","Mgp","Col3a1","Bgn","Dcn","Serpine1","Mmp9","Snai1","Egfl7"))
Seu.EC_all<-AddModuleScore(object = Seu.EC_all, features = mesenchymal_genes_for_score, name = "mesenchymal_score")

FeaturePlot(Seu.EC_all, features = "mesenchymal_score1", label = T, label.size = 4, split.by = "condition")
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/FeaturePlot_MesenchymalScore_ec_all.pdf"), width = 8, height = 5,  device = "pdf")
DO.BarplotWilcox(Seu.EC_all, Feature = "mesenchymal_score1", ctrl.condition = "Wildtype")
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/vlnplots/Barplot_MesenchymalScore_ec_all_wilcox.pdf"), width = 5, height = 5,  device = "pdf")
```

#### 6.2.1.3 Split cells based on Mesenchmyal score
```{r}
df <- FetchData(Seu.EC_all, vars = c("mesenchymal_score1", "condition"))

p<-ggplot(df, aes(x = mesenchymal_score1, color = condition, fill = condition)) +
  geom_density(alpha = 0.25, adjust = 1) +
  theme_classic()
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/Density_MesenchymalScore_ec_all.pdf"), width = 4, height = 2,  device = "pdf")
```


```{r}
df <- FetchData(Seu.EC_all, vars = c("mesenchymal_score1", "condition", "orig.ident")) %>%
  as.data.frame() %>%
  mutate(
    score_quantile = ntile(mesenchymal_score1, 10),
    score_quantile = factor(paste0("Q", score_quantile), levels = paste0("Q", 1:10))
  )

prop_df <- df %>%
  count(score_quantile, condition) %>%
  group_by(score_quantile) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

ggplot(prop_df, aes(x = score_quantile, y = prop, fill = condition)) +
  geom_col(position = "stack", color = "black", width = 0.8) +
  geom_text(
    aes(label = percent(prop, accuracy = 1)),
    position = position_stack(vjust = 0.5),
    size = 3
  ) +
  scale_y_continuous(labels = percent_format()) +
  theme_classic() +
  labs(x = "Quantile", y = "Proportion", fill = "Condition")
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/barplot_proportion_quant10_mesenchymalscore_ec_all.pdf"), width = 4, height = 2,  device = "pdf")
```


```{r}
library(Seurat)
library(dplyr)
library(ggplot2)
library(hrbrthemes)
library(viridis)

df <- FetchData(Seu.EC_all, vars = c("mesenchymal_score1", "condition", "orig.ident")) %>%
  as.data.frame() %>%
  mutate(
    score_quantile = ntile(mesenchymal_score1, 10),
    score_quantile = factor(paste0("Q", score_quantile), levels = paste0("Q", 1:10))
  )

prop_df <- df %>%
  count(score_quantile, condition, orig.ident) %>%
  group_by(score_quantile) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

ggplot(prop_df, aes(x = score_quantile, y = prop, fill = condition)) +
  geom_boxplot() +
  geom_jitter(color="black", size=1.4, alpha=0.9) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(x = "Mesenchymal score quantile", y = "Proportion of cells per quantile")
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/boxplot_mesenchymalscore_quant10_ec_all_combined.pdf"), width = 5, height = 5,  device = "pdf")
  

```


```{r}
df <- FetchData(Seu.EC_all, vars = c("mesenchymal_score1", "condition", "orig.ident")) %>%
  as.data.frame() %>%
  mutate(
    score_quantile = ntile(mesenchymal_score1, 4),
    score_quantile = factor(paste0("Q", score_quantile), levels = paste0("Q", 1:4))
  )

prop_df <- df %>%
  count(score_quantile, condition, orig.ident) %>%
  group_by(score_quantile) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()
prop_df$condition<-factor(prop_df$condition, levels = c("Wildtype","Tcf15mutant"))


ggplot(prop_df, aes(x = score_quantile, y = prop, fill = condition)) +
  geom_boxplot() +
  geom_point(position = position_jitterdodge()) +
  theme_classic() +
  scale_y_continuous(labels = scales::percent_format()) + 
  labs(x = "Mesenchymal score quantile", y = "Proportion of cells per quantile")
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/boxplot_mesenchymalscore_quant4_ec_all_combined.pdf"), width = 5, height = 5,  device = "pdf")

Seu.EC_all<-AddMetaData(object = Seu.EC_all, metadata = data.frame(quant4=df$score_quantile, row.names = rownames(df)))
p<-DO.CellComposition(Seu.EC_all, cluster_column = "quant4", sample_column = "orig.ident", condition_column = "condition", scanpro_plots = TRUE, outputFolder = paste0(outputFolder, "analysis/plots/celltype/EC_all/"))
p
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/composition_mesenchymalscore_quant4_ec_all.pdf"), width = 5, height = 8,  device = "pdf")
```


)

```{r}
hist(Seu.EC_all$mesenchymal_score1, breaks = 2000)

Seu.EC_all$mesenchymal_group <- ifelse(Seu.EC_all$mesenchymal_score1 > 0, "High", "Low")
p<-DO.CellComposition(Seu.EC_all, cluster_column = "mesenchymal_group", sample_column = "orig.ident", sort_x = c("Wildtype", "Tcf15mutant"))
p<- ggpubr::annotate_figure(p, bottom = "Mesenchymal Score > 0 = High; Mesenchymal Score < 0 = Low")
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/Barplot_CellComposiotion_MesenchymalGroup_ec_all_combined.pdf"), width = 5, height = 5,  device = "pdf")


#for individual celltypes
Seu.EC_all$condition_mesenchymal_group<- paste(Seu.EC_all$condition, Seu.EC_all$mesenchymal_group, sep = "_")
p<-DO.CellComposition(Seu.EC_all, cluster_column = "annotation",condition_column = "condition_mesenchymal_group", sample_column = "orig.ident")
p<- ggpubr::annotate_figure(p, bottom = "Mesenchymal Score > 0 = High; Mesenchymal Score < 0 = Low")
ggsave(filename = paste0(outputFolder, "analysis/plots/celltype/EC_all/Barplot_CellComposiotion_MesenchymalGroup_ec_all.pdf"), width = 10, height =4,  device = "pdf")

```


### 6.2.2 Capilary ECs
```{r}
table(SeuratObject.combined$annotation_refined)
Seu.EC_cap<-subset(SeuratObject.combined, subset = annotation_refined == "Cap_EC")

genes<-c("Serpine1","Egfl7","Pecam1", "Twist1", "Snai1")
for (gene in genes) {
  DO.Vln.Plot.wilcox(Seu.EC_cap, SeuV5 = T, Feature = gene, ctrl.condition = "Wildtype", group.by = "condition")
    filename<-paste0()
  ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_cap/VLNPlot_",gene,"_EC_cap.pdf"), width = 5, height = 5)
  
  DO.Mean.SEM.Graphs.cluster.t(Seu.EC_cap, Feature = gene, ctrl.condition = "Wildtype", group.by = "condition")
  filename<-paste0()
  ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_cap/Mean.SEM.cluster.tTest_",gene,"_EC_cap.pdf"), width = 5, height = 5)
  
  DO.Mean.SEM.Graphs.wilcox(Seu.EC_cap, Feature = gene, ctrl.condition = "Wildtype", group.by = "condition")
  filename<-paste0()
  ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_cap/Mean.SEM.wilcoxon_",gene,"_EC_cap.pdf"), width = 5, height = 5)
}


```

```{r}
mesenchymal_genes<-c("Mmp14","Fn1","Postn","Eln","Vin","Col1a1","Col1a2","Mgp","Col3a1","Bgn","Dcn","Tnc","Serpine1","Mmp9","Snai1","Snai2","Twist1","Zeb2","Il6","Pdgfra")

vascular_genes<-c("Pecam1","Kdr","Ptprb","Spry1","Foxo1","S1pr1","Efna1","Gja4","Cxcr4","Ace","Tgfbr2","Irp5")


p<- DO.Dotplot(Seu_object = Seu.EC_cap,
           Feature = mesenchymal_genes,
           group.by.x = "condition",
           plot.margin = c(1,1,1,1),
           annotation_x = F, # only useable if Feature is a dataframe with cluster information, e.g. FindAllMakers output (you need a column with cluster)
           point_stroke=0.1,
           annotation_x_rev = F,
           textSize = 14,
           hjust = 0.5,
           vjust = 0,
           textRot = 0,
           segWidth = 0.3,
           lwd = 3
        ) + ggtitle("Mesenchymal Genes in cap ECs")

ggsave(paste0(outputFolder, "analysis/plots/celltype/EC_all/dotplot_mesenchymalmarkers_ec_all.pdf"), width = 10, height = 5)

```





# 7. CellChat analysis

## 7.1 Create CellChat object for Empty

```{r}
library(CellChat)
library(patchwork)

Seu_Tcf15 <- subset(SeuratObject.combined, subset = condition =="Tcf15mutant")

#CellChat assumes a column called samples with orig.ident information
Seu_Tcf15$samples <- Seu_Tcf15$orig.ident
Seu_Tcf15$samples <- factor(Seu_Tcf15$samples)

CC_Tcf15 <- createCellChat(Seu_Tcf15,
                           group.by = "annotation_refined",
                           assay = "RNA")

#Initialize CellchatDB
CellChatDB <- CellChatDB.mouse
CC_Tcf15@DB <- CellChatDB

CC_Tcf15 <- subsetData(CC_Tcf15)
CC_Tcf15 <- identifyOverExpressedGenes(CC_Tcf15)
CC_Tcf15 <- identifyOverExpressedInteractions(CC_Tcf15)

#Try their projection approach to reduce dropout effects
CC_Tcf15 <- CellChat::smoothData(CC_Tcf15, adj = PPI.mouse)

#Comunication propability
CC_Tcf15 <- computeCommunProb(CC_Tcf15, population.size = T) # I will set this parameter to true since population sizes differ a lot
CC_Tcf15 <- computeCommunProbPathway(CC_Tcf15)

#Aggregated cell-cell communication
CC_Tcf15 <- aggregateNet(CC_Tcf15)
CC_Tcf15 <- netAnalysis_computeCentrality(CC_Tcf15)

```

## 7.2 Create CellChat object for ctrl

```{r}
Seu_Wildtype <- subset(SeuratObject.combined, subset = condition =="Wildtype")

#CellChat assumes a column called samples with orig.ident information
Seu_Wildtype$samples <- Seu_Wildtype$orig.ident
Seu_Wildtype$samples <- factor(Seu_Wildtype$samples)

CC_Wildtype <- createCellChat(Seu_Wildtype,
                           group.by = "annotation_refined",
                           assay = "RNA")

#Initialize CellchatDB
CellChatDB <- CellChatDB.mouse
CC_Wildtype@DB <- CellChatDB

CC_Wildtype <- subsetData(CC_Wildtype)
CC_Wildtype <- identifyOverExpressedGenes(CC_Wildtype)
CC_Wildtype <- identifyOverExpressedInteractions(CC_Wildtype)

#Try their projection approach to reduce dropout effects
CC_Wildtype <- CellChat::smoothData(CC_Wildtype, adj = PPI.mouse)

#Communication propability
CC_Wildtype <- computeCommunProb(CC_Wildtype, population.size = T) # I will set this parameter to true since population sizes differ a lot
CC_Wildtype <- computeCommunProbPathway(CC_Wildtype)

#Aggregated cell-cell communication
CC_Wildtype <- aggregateNet(CC_Wildtype)
CC_Wildtype <- netAnalysis_computeCentrality(CC_Wildtype)
```

## 7.3 Combined CellChat analysis

```{r}
#Continue with merging this object with the ctrl
CC_list <- list(wt = CC_Wildtype, tcf15 = CC_Tcf15)
CC_merge <- mergeCellChat(CC_list, add.names = names(CC_list))

saveRDS(CC_merge, file = paste0(outputFolder, "analysis/cellchat/TCF15_CC.rds"))
CC_merge <- readRDS(file = paste0(outputFolder, "analysis/cellchat/TCF15_CC.rds"))


pdf(paste0(outputFolder, "analysis/cellchat/InteractionNumber_all.pdf"))
netVisual_diffInteraction(CC_merge,
                          weight.scale = T,
                          # sources.use = 12,
                          title.name = "Outgoing Interaction All",
                          arrow.width = 1,
                          arrow.size = .5)
dev.off()

pdf(paste0(outputFolder, "analysis/cellchat/InteractionWeigth_all.pdf"))
netVisual_diffInteraction(CC_merge,
                          weight.scale = T,
                          # sources.use = 12,
                          title.name = "Interaction weight All",
                          arrow.width = 1,
                          arrow.size = .5,
                          measure = "weight")
dev.off()


#Number of interactions, weight
gg1 <- compareInteractions(CC_merge, show.legend = F, group = c(1,2), color.use = DOtools:::umap_colors)
gg2 <- compareInteractions(CC_merge, show.legend = F, group = c(1,2), measure = "weight", color.use = DOtools:::umap_colors)
gg1+gg2
ggsave(paste0(outputFolder, "analysis/cellchat/CompareInteractions_bar_plot.pdf"), width = 5, height = 3)


 #number of interaction per cell type
pdf(paste0(outputFolder, "analysis/cellchat/CompareInteractions_circ.pdf"), width = 14, height = 7)
par(mfrow = c(1,2), xpd=TRUE)
gg1 <- netVisual_diffInteraction(CC_merge, weight.scale = T, arrow.size = 0.05)
gg2 <- netVisual_diffInteraction(CC_merge, weight.scale = T, arrow.size = 0.05, measure = "weight")
dev.off()

#as heatmap
pdf(paste0(outputFolder, "analysis/cellchat/CompareInteractions_heatmap.pdf"), width = 8, height = 5)
par(mfrow = c(1,2), xpd=TRUE)
gg1 <- netVisual_heatmap(CC_merge)
gg2 <- netVisual_heatmap(CC_merge, measure = "weight")
gg1+gg2
dev.off()

#Identify up and downregulated signaling ligand-receptor pairs 
#sources 
levels(CC_merge@idents$joint)
for (src_name in levels(CC_merge@idents$joint)) {
  target_names <- paste(levels(CC_merge@idents$joint)[levels(CC_merge@idents$joint) != src_name], collapse = "-")
  print(src_name)
  p <- netVisual_bubble(CC_merge,
                   sources.use = grep(src_name,levels(CC_merge@idents$joint), fixed = T),
                   targets.use = which(levels(CC_merge@idents$joint) != src_name),
                   comparison = c(1,2),
                   angle.x = 45,
                   color.text = DOtools:::umap_colors)
  
  build_p <- ggplot_build(p)
  plot_names <- build_p$layout$panel_params[[1]]$y$get_labels()
  plot_names <- plot_names[!grepl("Ptprm  - Ptprm",plot_names)]
  
  pairLR_all <- CellChatDB.mouse$interaction
  pairLR_use <- pairLR_all[pairLR_all$interaction_name_2 %in% plot_names,]$pathway_name

  netVisual_bubble(CC_merge,
                   sources.use = grep(src_name,levels(CC_merge@idents$joint), fixed = T),
                   targets.use = which(levels(CC_merge@idents$joint) != src_name),
                   comparison = c(1,2),
                   angle.x = 45,
                   color.text = DOtools:::umap_colors,
                   signaling = pairLR_use)
  
 ggsave(paste0(outputFolder, "analysis/cellchat/LigandReceptor_analysis_",src_name,"_to_others.pdf"), width = 10, height = 17) 
 
}

for (src_name in levels(CC_merge@idents$joint)) {
  #create incoming-outgoing scatter plot with defined ident between conditions
  gg1 <- netAnalysis_signalingChanges_scatter(CC_merge,
                                              idents.use = src_name,
                                              color.use = DOtools:::umap_colors,
                                              top.label = 1,
                                              signaling.exclude = c("COLLAGEN","LAMININ","PTPRM","PECAM1"),
                                              comparison = c(1,2))
  gg1 <- gg1+
    DOtools:::theme_box()+
    labs(title = paste0("Signaling changes upon miRNAko in ", src_name))+
    theme(axis.text.x = element_text(color = "black",angle = 0,hjust = 1, size = 14),
            axis.text.y = element_text(color = "black", size = 14),
            axis.title.x = element_text(color = "black", size = 14),
            axis.title = element_text(size = 14, color = "black"),
            plot.title = element_text(size = 14, hjust = 0.5),
            axis.line = element_line(color = "black"),
            strip.text.x = element_text(size = 14, color = "black"),
            legend.position = "right",
            legend.text = element_text(size = 10, color = "black", family = "Helvetica"),
            legend.title = element_text(size = 10, color = "black", family = "Helvetica", hjust =0))+
    guides(color = ggplot2::guide_legend(title = "Specifity", order = 1),
           shape = ggplot2::guide_legend(title = "Direction", order = 2))
  gg1
  ggsave(paste0(outputFolder, "analysis/cellchat/Signaling_changes_in_", src_name,"_upon_tcf15ko_LAMININ_COLLAGEN_PTPRM_PECAM1_excluded.pdf"), width = 8, height = 7)
}




```

## 7.4 Iteraction Circos only Ec's and FB clusters
```{r}
pdf(paste0(outputFolder, "analysis/cellchat/InteractionNumber_FB.pdf"))
netVisual_diffInteraction(CC_merge,
                          weight.scale = T,
                          sources.use = c(1,11),
                          title.name = "Outgoing Interaction FB",
                          arrow.width = 1,
                          arrow.size = .5)
dev.off()

pdf(paste0(outputFolder, "analysis/cellchat/InteractionWeigth_FB.pdf"))
netVisual_diffInteraction(CC_merge,
                          weight.scale = T,
                          sources.use = c(1,11),
                            title.name = "Interaction weight FB",
                          arrow.width = 1,
                          arrow.size = .5,
                          measure = "weight")
dev.off()

```


```{r}
pdf(paste0(outputFolder, "analysis/cellchat/InteractionNumber_EC.pdf"))
netVisual_diffInteraction(CC_merge,
                          weight.scale = T,
                          sources.use = c(4),
                          title.name = "Outgoing Interaction EC",
                          arrow.width = 1,
                          arrow.size = .5)
dev.off()

pdf(paste0(outputFolder, "analysis/cellchat/InteractionWeigth_EC.pdf"))
netVisual_diffInteraction(CC_merge,
                          weight.scale = T,
                          sources.use = c(2,4,8,9,12),
                          title.name = "Interaction weight EC",
                          arrow.width = 1,
                          arrow.size = .5,
                          measure = "weight")
dev.off()
```



### 7.5 Bubble plot of CapEC's to selected celltypes
```{r}

#ECs to FB
  pairLR_all <- CellChatDB.mouse$interaction
  pairLR_use <- pairLR_all[pairLR_all$interaction_name_2 %in% plot_names,]$pathway_name
  
  p<-netVisual_bubble(CC_merge,
                   sources.use = c("Cap_EC", "End_EC","EC6_ven", "Art_EC"),
                   targets.use = c("FB"),
                   comparison = c(1,2),
                   angle.x = 45,
                   color.text = DOtools:::umap_colors)
  
ggsave(paste0(outputFolder, "analysis/cellchat/LigandReceptor_analysis_AllEcs_to_FB.pdf"), width = 4, height = 3)

#FB to EC's
  pairLR_all <- CellChatDB.mouse$interaction
  pairLR_use <- pairLR_all[pairLR_all$interaction_name_2 %in% plot_names,]$pathway_name
  
  p<-netVisual_bubble(CC_merge,
                   sources.use = c("FB"),
                   targets.use = c("Cap_EC", "End_EC","EC6_ven","FB"),
                   comparison = c(1,2),
                   angle.x = 45,
                   signaling = c()
                   color.text = DOtools:::umap_colors)
  
ggsave(paste0(outputFolder, "analysis/cellchat/LigandReceptor_analysis_FB_to_AllEcs+FB.pdf"), width = 4, height = 7)
```

## 7.4 Circos plot ArtEC vs rest & CapEC vs rest

```{r}
par(mfrow = c(1,2), xpd=T)
levels(CC_merge@idents$joint)

for (i in 1:length(CC_list)) {
  netVisual_chord_gene(CC_list[[i]],
                       sources.use = 12,
                       targets.use = setdiff(1:length(levels(CC_merge@idents$joint)),12),
                       lab.cex = 0.5, title.name = paste0("Signaling from ArtEC - ", names(CC_list)[i]))
}
```

## 7.5 Output Data

```{r}
int_all <- list()
for (celltype in unique(c(CC_list$tcf15@idents))) {
  for (cellchat in CC_list){
    cellchat <- rankNetPairwise(cellchat)
    for (i in unique(cellchat@meta$annotation_ref)){
      name <- paste('from', i,'to',celltype, 'in', unique(cellchat@meta$condition))
      int <- identifyEnrichedInteractions(cellchat, from = i, to=celltype, bidirection=F)
      if(nrow(int) != 0){
        int$name = name
        int$sender = i
        int$receiver = celltype
        print (cellchat@meta$condition)
        int$condition = unique(cellchat@meta$condition)
        int$value <- cellchat@net$prob[celltype, i, rownames(int)]
        int_all[[name]]<-int
      }
    }
    
    for (j in as.character(unique(cellchat@meta$annotation_ref))[!as.character(unique(cellchat@meta$annotation_ref))==celltype]){
      name <- paste('from',celltype,  'to', j, 'in', unique(cellchat@meta$condition))
      if(nrow(int) != 0){
        int$name = name
        int$sender = celltype
        int$receiver = j
        int$condition = unique(cellchat@meta$condition)
        int$value <- cellchat@net$prob[j, celltype, rownames(int)]
        int_all[[name]]<-int
      }  
    }
  }
}
Interaction.db <- purrr::reduce(int_all, full_join)

wb <- createWorkbook()
hs1 <- createStyle(
  fgFill = '#DCE6F1', halign='CENTER', textDecoration='bold', border='Bottom'
  )

addWorksheet(wb, 'Interactions_all')
writeData(wb, sheet='Interactions_all', x=Interaction.db, rowNames=T, headerStyle=hs1)
saveWorkbook(wb, file=paste0(outputFolder,'Cellchat_interaction_ALL.xlsx'))


```

```{r}
SeuratObject.combined$orig.ident<-factor(SeuratObject.combined$orig.ident, levels=c("Wildtype-1","Wildtype-2","Wildtype-3","Tcf15mutant-1","Tcf15mutant-2","Tcf15mutant-3"))
p<-DO.Dotplot(object = SeuratObject.combined, Feature = "Tcf15", group.by.x = "orig.ident", group.by.y = "predicted_labels_celltypist")
ggsave(filename = paste0(outputFolder, "analysis/plots/global/vlnplots/Barplot_Tcf15_global.pdf"), width = 10, height = 15, plot = p, device = "pdf")

```
