---
title: "Analysis for Elvin TCF15KO"
author: "Mariano Ruz Jurado"
date: "2024-08-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

-##########-
# -) START DEG analysis with DESEQ2

##0. Load libraries, set sources
```{r}
library(ggplot2)
library(ggfortify)
library(gtools)
library(tidyverse)
library(reshape2)
library(tibble)
library(EnhancedVolcano)
library(openxlsx)
```

##1. Load in GTF and adjust it fo further analysis
```{r}
gtf <- data.table::fread("/mnt/mariano/Genomes_Annotations/human/GRCH38/Homo_sapiens.GRCh38.77.gtf", sep = "\t", blank.lines.skip = T)
data.table::setnames(gtf, names(gtf), c("chr","source","type", "start", "end", "score", "strand", "phase", "attributes"))

gtf<-gtf[type=="gene"]

extract_attributes <- function(gtf_attributes, att_of_interest){
  
  att <- strsplit(gtf_attributes, "; ")
  att <- gsub("\"", "", unlist(att))
  if(!is.null(unlist(strsplit(att[grep(att_of_interest, att)], " ")))){
    return(unlist(strsplit(att[grep(att_of_interest, att)], " "))[2])
  }else{
      return(NA)
    }
}
  
gtf$biotype <- unlist(lapply(gtf$attributes, extract_attributes, "gene_biotype"))
gtf$gene_id <- unlist(lapply(gtf$attributes, extract_attributes, "gene_id"))
```

##2. FPKM analysis
```{r}
fpkms <- read.table(file = "/mnt/mariano/scStorage/Mariano/forElvin/24063/cuffdiff/genes.read_group_tracking", header = T)
fpkms <- data.frame(gene=fpkms$tracking_id, condition=fpkms$condition, replicate=fpkms$replicate, fpkm=fpkms$FPKM)

fpkm_unmelt <- reshape2::dcast(fpkms, formula = gene~condition+replicate, fill = "fpkm")


#read DEG statistic
stat<-read.table(file = "/mnt/mariano/scStorage/Mariano/forElvin/24063/cuffdiff/gene_exp.diff", header = T)
stat$test_id<-NULL
combinations<-mixedsort(unique(paste(stat$sample_1, stat$sample_2, sep="|")))

outTable<-fpkm_unmelt
for (comb in combinations) {
  comb<-unlist(strsplit(comb, split = "|", fixed = T))
  condTable<-subset(stat,sample_1==comb[1] & sample_2==comb[2])
  colN<-colnames(condTable)
  cond=paste(condTable[1,5],condTable[1,4],sep = "/")
  colnames(condTable)[7:13]=c(paste("Mean",condTable[1,4],sep="_"),
                              paste("Mean",condTable[1,5],sep="_"),
                              paste(colN[9],cond,sep="_"),
                              paste(colN[10],cond,sep="_"),
                              paste(colN[11],cond,sep="_"),
                              paste(colN[12],cond,sep="_"),
                              paste(colN[13],cond,sep="_"))
  outTable <- cbind(outTable,condTable[,7:13])
}

outTable2<-add_column(outTable, gene_id=condTable$gene, .after=1, .name_repair = "minimal")
outTable2$Treshold_above_1 <- as.factor(ifelse(outTable2$Mean_Control > 1|outTable2$Mean_TCF15KO > 1, ">1", "<1"))

data_df <- outTable2[which(outTable2$Treshold_above_1 == ">1"),3:10] # exchange number corresponding to your samples

# Identify non-numeric columns
non_numeric_cols <- sapply(data_df, function(x) !is.numeric(x))

# Convert non-numeric values to NA
data_df[non_numeric_cols] <- lapply(data_df[non_numeric_cols], function(x) {
  as.numeric(as.character(x))
})


pca_result <- prcomp(t(data_df), scale. = T)
pca_df <- as.data.frame(pca_result$x)

# PCA result is stored in pca_result
pca_df <- as.data.frame(pca_result$x)

# Subset the columns for PCA1, PCA2
pca_subset <- pca_df[, c("PC1", "PC2")]

# Calculate variance explained
var_explained <- (pca_result$sdev^2) / sum(pca_result$sdev^2)

# Create a data frame for groups
pca_subset$groups <- c(rep("Control", 4), rep("TCF15KO", 4))
plot(pca_result$x)



# Plot the panel of PCA plots with colorized points and variance explained
p<-ggplot(pca_subset, aes(x = PC1, y = PC2, color = groups)) +
  geom_point(size = 4) +
  scale_color_manual(values = c("tomato", "royalblue", "forestgreen","gold", "darkred", "darkblue")) + ggtitle("PCA Control, TCF15_OE") +
  theme(plot.title = element_text(hjust = 0.5))
p
ggsave(filename = "/media/Helios_scStorage/Mariano/forElvin/24063/cuffdiff/PCA_CTRL_TCF15KO.pdf", plot = p, width = 6, height = 4)
```


##3. Write DEG lists
```{r}
logcolnames <- grep(c("log2"), colnames(outTable2), value = TRUE)
logcolnames <- grep(c("Control"), logcolnames, value = TRUE)

for (clnames in logcolnames) {
  conditions_rel <- unlist(strsplit(clnames, "_"))[3]
  conditions_rel_split <- unlist(strsplit(conditions_rel,  "/"))
  conditions_rel_split <- c("gene", conditions_rel_split) # include gene column
  
  outTable3 <- outTable2[!is.infinite(outTable2[[clnames]]),]
  outTable3 <- subset(outTable3, outTable3$Treshold_above_1==">1")
  
  #remove duplicate gene names
  outTable3 <- outTable3[!duplicated(outTable3$gene_id),]
  gtf2<-data.frame(gene=gtf$gene_id, biotype=gtf$biotype)
  outTable4 <- merge(outTable3, gtf2, by = "gene")
  
  result <- Reduce(cbind, lapply(conditions_rel_split, function(condition) {
    outTable4[, grep(condition, colnames(outTable4), value = TRUE)]
  }))
  
  # Extract relevant entries from conditions_rel_split
  gene_pattern <- conditions_rel_split[1]
  condition1 <- conditions_rel_split[2]
  condition2 <- conditions_rel_split[3]

  # Create a dynamic regex pattern to include only Mean columns without numbers
  pattern <- paste0("(", gene_pattern, "|", condition1, ".*", condition2, "|", 
                    condition2, ".*", condition1, "|", condition1, "(_|Mean$)", "|", 
                    condition2, "(_|Mean$)", "|^Mean_", condition1, "$|^Mean_", condition2, "$)")
  
  # Filter the columns using grep
  filtered_result <- result[, grep(pattern, colnames(result), value = TRUE)]
  # Remove columns that have a period followed by a number
  filtered_result <- filtered_result[, !grepl("\\.[0-9]+$", colnames(filtered_result))]
  
  #write Excel sheets
  openxlsx::write.xlsx(filtered_result, file = paste0("/mnt/mariano/scStorage/Mariano/forElvin/24063/cuffdiff/TCF15KO_bulk_seq_human_24063_",condition1,"_",condition2,"_noInf_thresh>1.xlsx"), colNames=T, rowNames=F)
}
```


# -) END DEG analysis with DESEQ2
-##########-

# -) Figure 2b

##4. Request_250514 Additional Analysis: Combine GO Terms, provide list of genes
```{r}
DEG_DF <- read.xlsx("/mnt/mariano/NGS_storage/Elvin/24063/cuffdiff/TCF15KO_bulk_seq_human_24063_TCF15KO_Control_noInf_thresh>1.xlsx", colNames=T, rowNames=F)

DEG_DF_sig <- DEG_DF[DEG_DF$`significant_TCF15KO/Control` == "yes",] #only significant

DEG_DF_sig <- DEG_DF_sig[abs(DEG_DF_sig$`log2.fold_change._TCF15KO/Control`) >= 0.25, ] # thresh for FC

#divide by pos and negative regulated genes
positive_DEG_DF <- DEG_DF_sig[DEG_DF_sig$`log2.fold_change._TCF15KO/Control` > 0,]
negative_DEG_DF <- DEG_DF_sig[DEG_DF_sig$`log2.fold_change._TCF15KO/Control` < 0,]

#read in excel sheet results from metascape
#up
GO.UP <- openxlsx::read.xlsx("/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/DEG_UP/metascape_result/metascape_result.up.xlsx",sheet = 2)
#down
GO.DOWN <- openxlsx::read.xlsx("/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/DEG_DOWN/metascape_result/metascape_result.down.xlsx",sheet = 2)

#Just summary terms for plot
GO.UP.Sum <- GO.UP[grepl("Summary",GO.UP$GroupID),]
GO.DOWN.Sum <- GO.DOWN[grepl("Summary",GO.DOWN$GroupID),]


##For GO.UP

#Combine the two rows VEGFA VEGFR2 signaling ith vasculature development
rows_to_merge <- grepl("VEGFA VEGFR2 signaling", GO.UP.Sum$Description) | grepl("vasculature development", GO.UP.Sum$Description)
to_merge <- GO.UP.Sum[rows_to_merge, ]

#Create the combined row
combined_row <- to_merge[1, ]  # start from the first row as template
combined_row$Description <- "VEGFA VEGFR2 signaling_vasculature development"
combined_row$Genes <- paste(to_merge$Genes, collapse = ",")
combined_row$Symbols <- paste(to_merge$Symbols, collapse = ",")

# Remove the original rows and add the new one
GO.UP.Sum <- GO.UP.Sum[!rows_to_merge, ]
GO.UP.Sum <- rbind(GO.UP.Sum, combined_row)

##For GO.DOWN

#Combine the two rows VEGFA VEGFR2 signaling ith vasculature development
rows_to_merge <- grepl("VEGFA VEGFR2 signaling", GO.DOWN.Sum$Description) | grepl("vasculature development", GO.DOWN.Sum$Description)
to_merge <- GO.DOWN.Sum[rows_to_merge, ]

#Create the combined row
combined_row <- to_merge[1, ]  # start from the first row as template
combined_row$Description <- "VEGFA VEGFR2 signaling_vasculature development"
combined_row$Genes <- paste(to_merge$Genes, collapse = ",")
combined_row$Symbols <- paste(to_merge$Symbols, collapse = ",")

# Remove the original rows and add the new one
GO.DOWN.Sum <- GO.DOWN.Sum[!rows_to_merge, ]
GO.DOWN.Sum <- rbind(GO.DOWN.Sum, combined_row)


#Take *-1 to the log10p column, everything positive
GO.UP.Sum$LogP <- GO.UP.Sum$LogP*(-1)
GO.UP.Sum$`Log(q-value)` <- GO.UP.Sum$`Log(q-value)`*(-1)
GO.DOWN.Sum$LogP <- GO.DOWN.Sum$LogP*(-1)
GO.DOWN.Sum$`Log(q-value)` <- GO.DOWN.Sum$`Log(q-value)`*(-1)

#Add column with avg FC UP
GO.UP.Sum$avg_FC <- NA
for (groups in GO.UP.Sum$GroupID) {
  relev_genes <- unlist(strsplit(GO.UP.Sum[GO.UP.Sum$GroupID == groups,]$Symbols,split = ","))
  GO.UP.Sum[GO.UP.Sum$GroupID == groups,]$avg_FC <- mean(DEG_DF_sig[DEG_DF_sig$gene_id %in% relev_genes,]$`log2.fold_change._TCF15KO/Control`)
}

#Add column with avg FC UP
GO.DOWN.Sum$avg_FC <- NA
for (groups in GO.DOWN.Sum$GroupID) {
  relev_genes <- unlist(strsplit(GO.DOWN.Sum[GO.DOWN.Sum$GroupID == groups,]$Symbols,split = ","))
  GO.DOWN.Sum[GO.DOWN.Sum$GroupID == groups,]$avg_FC <- mean(DEG_DF_sig[DEG_DF_sig$gene_id %in% relev_genes,]$`log2.fold_change._TCF15KO/Control`)
}

#Comb DF by relevant columns
GO_TCF15down_Subset <- GO.DOWN.Sum[, c("Description", "Log(q-value)", "avg_FC")]
GO_TCF15Up_Subset <- GO.UP.Sum[, c("Description", "Log(q-value)", "avg_FC")]

#Add p-values and terms which were also found in the other condition, otherwise dont add them so that in the plot they are just empty spaces
terms_TCF15 <- GO_TCF15Up_Subset$Description
GO.DOWN_TCF15 <- GO.DOWN[GO.DOWN$Description %in% terms_TCF15,]
GO.DOWN_TCF15$avg_FC <- NA
for (groups in GO.DOWN_TCF15$GroupID) {
  relev_genes <- unlist(strsplit(GO.DOWN_TCF15[GO.DOWN_TCF15$GroupID == groups,]$Symbols,split = ","))
  GO.DOWN_TCF15[GO.DOWN_TCF15$GroupID == groups,]$avg_FC <- mean(DEG_DF_sig[DEG_DF_sig$gene_id %in% relev_genes,]$`log2.fold_change._TCF15KO/Control`)
}
GO.DOWN_TCF15 <- GO.DOWN_TCF15[, c("Description", "Log(q-value)", "avg_FC")]
GO.DOWN_TCF15$`Log(q-value)` <- GO.DOWN_TCF15$`Log(q-value)` *-1

#add the ones missing
GO.DOWN_TCF15_missing <- GO.DOWN_TCF15[!(GO.DOWN_TCF15$Description %in% GO_TCF15down_Subset$Description),]
GO_TCF15DOWN_final <- rbind(GO_TCF15down_Subset,GO.DOWN_TCF15_missing)

terms_CTRL <- GO_TCF15down_Subset$Description
GO.UP_CTRL <- GO.UP[GO.UP$Description %in% terms_CTRL,]
GO.UP_CTRL$avg_FC <- NA
for (groups in GO.UP_CTRL$GroupID) {
  relev_genes <- unlist(strsplit(GO.UP_CTRL[GO.UP_CTRL$GroupID == groups,]$Symbols,split = ","))
  GO.UP_CTRL[GO.UP_CTRL$GroupID == groups,]$avg_FC <- mean(DEG_DF_sig[DEG_DF_sig$gene_id %in% relev_genes,]$`log2.fold_change._TCF15KO/Control`)
}
GO.UP_CTRL <- GO.UP_CTRL[, c("Description", "Log(q-value)", "avg_FC")]
GO.UP_CTRL$`Log(q-value)` <- GO.UP_CTRL$`Log(q-value)` *-1

#add the ones missing
GO.UP_CTRL_missing <- GO.UP_CTRL[!(GO.UP_CTRL$Description %in% GO_TCF15Up_Subset$Description),]
GO_TCF15UP_final <- rbind(GO_TCF15Up_Subset,GO.UP_CTRL_missing)


#Add information about in which group relevant
GO_TCF15DOWN_final$Group <- "DEGs Down"
GO_TCF15UP_final$Group <- "DEGs Up"

#Row-bind the subsets
GO_DF_Dotplot <- rbind(GO_TCF15UP_final, GO_TCF15DOWN_final)

#some re-naming
colnames(GO_DF_Dotplot) <- c("Term","p.adj","avg_FC","Group") # q-value is p-value adjusted by BH

#terms to 10
GO_DF_Dotplot <- GO_DF_Dotplot[order(GO_DF_Dotplot$p.adj, decreasing = T),]

GO_DF_Dotplot.final <- GO_DF_Dotplot %>%
  group_by(Group) %>%
  arrange(desc(p.adj)) %>%
  slice_head(n = 8) %>%
  ungroup()

GO_DF_Dotplot.final$Term <- factor(GO_DF_Dotplot.final$Term, levels = rev(c(GO_DF_Dotplot.final[GO_DF_Dotplot.final$Group == "DEGs Down",]$Term, GO_DF_Dotplot.final[GO_DF_Dotplot.final$Group == "DEGs Up" & GO_DF_Dotplot.final$Term != "VEGFA VEGFR2 signaling_vasculature development",]$Term)))

pmain <- ggplot2::ggplot(GO_DF_Dotplot.final, ggplot2::aes(x = Group,y = Term)) + ggplot2::theme_bw(base_size = 14) + 
    ggplot2::xlab("") + ggplot2::ylab("") + ggplot2::labs(title = "GO analysis TCF15KD")+ ggplot2::coord_fixed(clip = "off") + 
    ggplot2::theme(plot.margin = ggplot2::margin(t = 1, 
                                                 r = 0,
                                                 b = 1,
                                                 l = 0, 
                                                 unit = "cm"),
                   axis.text = ggplot2::element_text(color = "black"),
                   legend.direction = "horizontal",
                   axis.text.x = element_text(color = "black",angle = 45,hjust = 1, size = 9, family = "Helvetica"),
                   axis.text.y = element_text(color = "black", size = 9, family = "Helvetica"),
                   axis.title.x = element_text(color = "black", size = 10, family = "Helvetica"),
                   axis.title = element_text(size = 10, color = "black", family = "Helvetica"),
                   plot.title = element_text(size = 10, hjust = 0.5,face="bold", family = "Helvetica"),
                   plot.subtitle = element_text(size = 10, hjust = 0, family = "Helvetica"),
                   axis.line = element_line(color = "black"),
                   strip.text.x = element_text(size = 8, color = "black", family = "Helvetica"),
                   legend.text = element_text(size = 8, color = "black", family = "Helvetica"),
                   legend.title = element_text(size = 8, color = "black", family = "Helvetica", hjust =0),
                   legend.position = "right",
                   panel.grid.major = element_blank(),
                   panel.grid.minor = element_blank(),)
  
guides.layer <- ggplot2::guides(fill = ggplot2::guide_colorbar(title = "Avg.log2FC",
                                                               title.position = "top",
                                                               title.hjust = 0.5,
                                                               barwidth = unit(3.5,"cm"),
                                                               barheight = unit(0.5,"cm"),
                                                               frame.colour = "black",
                                                               frame.linewidth = 0.3,
                                                               ticks.colour = "black",
                                                               order = 2),
                                size = ggplot2::guide_legend(title = "-Log10p value", 
                                                             title.position = "top", title.hjust = 0.5, label.position = "bottom", 
                                                             override.aes = list(color = "black", fill = "grey50"), 
                                                             keywidth = ggplot2::unit(0.3, "cm"),
                                                             order = 1))

dot.col = c("#fff5f0","#990000") # TODO change the scalefillgradient to +n in the else part
gradient_colors <- c(
    "royalblue4",
    "#1b699e",
    "lightsteelblue",
    "#E1E1E9",
    "#E4E4EA",
    "#fff5f0",
    "#FFEEE4",
    "#FEC196",
    "#ea7e1e",
    "#ED5B41",
    "firebrick4")

fill.values <- GO_DF_Dotplot.final$avg_FC
size.value <- GO_DF_Dotplot.final$p.adj
breaks <- scales::breaks_extended(n=5)(range(fill.values))

if (abs(max(range(fill.values))) > abs(min(range(fill.values)))) {
  limits <- c(-max(range(fill.values)),max(range(fill.values))) 
} else{
  limits <- c(min(range(fill.values)),-min(range(fill.values))) 
}

breaks <- unique(sort(c(-breaks,breaks), decreasing = F))[seq(1, length(c(-breaks,breaks)), by=2)]
breaks <- breaks[!is.na(breaks)]

if (max(breaks) > max(limits)) {
  limits[length(limits)] <- breaks[length(breaks)]
}
if (min(breaks) < min(limits)) {
  limits[1] <- breaks[1]
}

pmain <- pmain + ggplot2::scale_fill_gradientn(colours = gradient_colors,
                                               breaks = breaks,
                                               limits = limits)


pmain <- pmain + 
  ggplot2::geom_point(ggplot2::aes(fill = fill.values,
                                   size = size.value,
                                   )
                      ,
                      shape = 21,
                      stroke=0.65) +
  guides.layer +
  ggplot2::scale_size(range = c(3,6)) +
  ggplot2::scale_size_continuous(breaks = pretty(round(as.vector(quantile(size.value))), n =10)[seq(1, 10, by = 2)],
                                 limits = c(min(size.value)*1.00,max(size.value)*1.05))+
  theme(panel.spacing = unit(0, "lines")) +
  ggplot2::scale_fill_gradientn(limits = c(-1,0.52), oob = scales::squish, colors = gradient_colors)
  
print(pmain)

ggsave("/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/GO_bubbleplot_TCF15KD_FC_p_adj_combterm.pdf",width = 5.5, height = 5.5)
write.xlsx(GO_DF_Dotplot.final, file = "/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/DF_GO_bubbleplot_TCF15KD_FC_p_adj_combterm.xlsx")

#save the genes in the combined term
row_comb_down <- GO.DOWN.Sum[GO.DOWN.Sum$Description == "VEGFA VEGFR2 signaling_vasculature development",]
write.xlsx(row_comb_down, file = "/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/VEGFA VEGFR2 signaling_vasculature development_term_DOWN_genes.xlsx")

row_comb_up <- GO.UP.Sum[GO.UP.Sum$Description == "VEGFA VEGFR2 signaling_vasculature development",]
write.xlsx(row_comb_up, file = "/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/VEGFA VEGFR2 signaling_vasculature development_term_UP_genes.xlsx")

```

# -) Figure 2c & Extended Data Figure 3f

##5. Volcano plot
```{r}
#
vasc_term_down <- read.xlsx("/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/VEGFA VEGFR2 signaling_vasculature development_term_DOWN_genes.xlsx")
vasc_term_up <- read.xlsx("/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/VEGFA VEGFR2 signaling_vasculature development_term_UP_genes.xlsx")

#VEGFA VEGFR2 signaling_vasculature development
vasc_term <- rbind(vasc_term_down,vasc_term_up)
vasc_term$Group <- c("DEGs Down", "DEGs Up")

#for elvin save the genes in the term
vasc_term_save <- vasc_term %>% 
  dplyr::select(Description, Group, Symbols)
write.xlsx(vasc_term_save, file = "/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/VEGFA VEGFR2 signaling_vasculature development_term_UP_DOWN_genes.xlsx")

#separate Gene Symbols by respecting the group
vasc_term_df <- vasc_term %>% 
  dplyr::select(Group, Symbols) %>%
  tidyr::separate_rows(Symbols, sep = ",")

vasc_term_df <- vasc_term_df[!duplicated(vasc_term_df$Symbols),]

#Add foldchange to the gene
DEG_DF <- read.xlsx("/mnt/mariano/NGS_storage/Elvin/24063/cuffdiff/TCF15KO_bulk_seq_human_24063_TCF15KO_Control_noInf_thresh>1.xlsx", colNames=T, rowNames=F)
DEG_DF_sig <- DEG_DF[DEG_DF$`significant_TCF15KO/Control` == "yes",] #only significant
DEG_DF_sig <- DEG_DF_sig[abs(DEG_DF_sig$`log2.fold_change._TCF15KO/Control`) >= 0.25, ] # thresh for FC

#Sort out genes which are not found in the RNA-seq
vasc_term_df <- vasc_term_df[vasc_term_df$Symbols %in% DEG_DF_sig$gene_id,]
#add fc
vasc_term_df$log2FC <- DEG_DF_sig[["log2.fold_change._TCF15KO/Control"]][match(vasc_term_df$Symbols,DEG_DF_sig$gene_id)]

vasc_term_df$Mean_TCF15KO <- DEG_DF_sig[["Mean_TCF15KO"]][match(vasc_term_df$Symbols,DEG_DF_sig$gene_id)]
vasc_term_df$Mean_Control <- DEG_DF_sig[["Mean_Control"]][match(vasc_term_df$Symbols,DEG_DF_sig$gene_id)]


DEG_DF$`q_value_TCF15KO/Control`[DEG_DF$`q_value_TCF15KO/Control` < 0.0005] <- 0.0005

p <- EnhancedVolcano(DEG_DF,
DEG_DF$gene_id,
selectLab = vasc_term_df[abs(vasc_term_df$log2FC) > 1,]$Symbols,
x = 'log2.fold_change._TCF15KO/Control',
y = 'q_value_TCF15KO/Control',
title = paste0("TCF15KD vs CTRL - Vasculature developing+VEGFA VEGFR2 signaling"),
subtitle = "",
pCutoff = 10e-2,
FCcutoff = 0.5,
pointSize = 1.5,
labSize = 3,
drawConnectors = T,
boxedLabels = F,
labFace = "bold",
#maxoverlapsConnectors = 6,
max.overlaps = Inf,
min.segment.length = 0.5,
typeConnectors = "closed",
widthConnectors = 0.25,
gridlines.major = F,
gridlines.minor = F,
lengthConnectors = unit(0.007, "npc"),
colConnectors = "darkgrey",
ylim = c(0,3.5),
xlim = c(-3.5, 3.5),
# colAlpha = 1/2,
col = c("grey50", "palegreen4", "#1b699e", "#ea6d1e")
)+   theme(legend.position = "none")
p


ggsave(filename = "/mnt/mariano/NGS_storage/Elvin/24063/GO_analysis/plots/volcano_TCF15KD_VEGFA VEGFR2 signaling_vasculature development.pdf",
       plot = p, device = "pdf", height = 8, width = 12)




```

#-) Extended Data Figure 1b

##6. Load AMI Timecourse new object
```{r}
Seu_AMI <- readRDS("/mnt/mariano/scStorage/DavidR/BioData/LukasT_TimeCourse/Objects/241206_LukasTAMI_scRNA_onlyYoung.rds")
DO.UMAP(Seu_AMI, group.by = "annotation")

#Tcf15 expression UMAP
DO.UMAP(Seu_AMI, features = "Tcf15",FeaturePlot = T, group.by = "annotation", repel=T)
ggsave(paste0(outputFolder,"/analysis/plots/UMAP_Tcf15_Expr.pdf"), width = 8, height = 8)


Seu_AMI_EC <- DO.Subset(Seu_AMI, ident = "annotation", ident_name = c("CapEC", "EndoEC", "VeinEC", "ArtEC"))
DO.UMAP(Seu_AMI_EC, features = "Tcf15",FeaturePlot = T, group.by = "annotation", repel=T)
ggsave(paste0(outputFolder,"/analysis/plots/UMAP_ECs_Tcf15_Expr.pdf"), width = 7, height = 7)

#Tcf15 Dotplot
DO.Dotplot(Seu_AMI_EC,
           Feature = "Tcf15",
           group.by.x = "annotation",
           coord_flip = F)

ggsave(paste0(outputFolder,"/analysis/plots/Dotplot_ECs_Tcf15_Expr.pdf"), width = 5, height = 5)


```

###6.1 Reclustering of the subsetted ECs

```{r}
#redo clustering steps
Seu_AMI_EC <- NormalizeData(object = Seu_AMI_EC)
Seu_AMI_EC <- ScaleData(object = Seu_AMI_EC)
Seu_AMI_EC <- FindNeighbors(object = Seu_AMI_EC, reduction = "integrated.harmony", dims = 1:10)
Seu_AMI_EC <- FindClusters(Seu_AMI_EC, resolution = 0.3, algorithm = 4, random.seed = 1234)
Seu_AMI_EC <- RunUMAP(object = Seu_AMI_EC, reduction = "integrated.harmony", dims = 1:10, spread = 5, min.dist = 0.1, reduction.name = "umap")

p1 <- DO.UMAP(Seu_AMI_EC, features = "Tcf15", group.by = "annotation", repel=T, label = T, legend.position = "none", umap_colors = c("#ff7f0e","#1f77b4","#c49c94","tomato2"))
p1
p2 <- DO.UMAP(Seu_AMI_EC, features = "Tcf15",FeaturePlot = T, group.by = "annotation", repel=T, min.cutoff="q10", max.cutoff="q90")
p3 <- DO.UMAP(Seu_AMI_EC, features = "Tcf15",DensityPlot = T, group.by = "annotation", repel=T)

ggpubr::ggarrange(plotlist = list(p1,p2,p3), ncol=3, nrow=1)
ggsave(paste0(outputFolder,"/analysis/plots/UMAP_Recluster_ECs_Tcf15_Expr.svg"), width = 12, height = 5)

DO.UMAP(Seu_AMI_EC, features = "Tcf15",FeaturePlot = T, group.by = "annotation", repel=T, min.cutoff="q10", max.cutoff="q90", legend.position = "right")
ggsave(paste0(outputFolder,"/analysis/plots/UMAP_Recluster_ECs_Tcf15_FeaturePlot.svg"), width = 6, height = 5)

DO.UMAP(Seu_AMI_EC, features = "Tcf15",DensityPlot = T, group.by = "annotation", repel=T, legend.position = "right")
ggsave(paste0(outputFolder,"/analysis/plots/UMAP_Recluster_ECs_Tcf15_DensityPlot.svg"), width = 6, height = 5)


```


#-) Extended Data Figure 1d

##7. UMAP CTRL.LV with TCF15 expression
```{r}
Human_HF_complete <- readRDS("/mnt/mariano/scStorage/Mariano/Human_Mice_Comparison/Seurat_and_R_objects/260402_Human_HF_complete.rds")

Human_HF_CTRLlv <- DO.Subset(Human_HF_complete, ident = "disease", ident_name = "CTRL.LV")
p1 <- DO.UMAP(Human_HF_CTRLlv, group.by = "annotation", raster=TRUE)
p2 <- DO.UMAP(Human_HF_CTRLlv, group.by = "annotation", DensityPlot = TRUE, features = "TCF15", raster = FALSE, label = TRUE)

ggpubr::ggarrange(plotlist = list(p1, p2), ncol = 2, nrow = 1)
ggsave(paste0(outputFolder,"/analysis/plots/UMAP_HCA_leftventricle_TCF15_Expr.pdf"), width = 12, height = 5)

DO.UMAP(Human_HF_CTRLlv, group.by = "annotation", DensityPlot = TRUE, features = "TCF15", raster = FALSE, label = TRUE, legend.position = "right")
ggsave(paste0(outputFolder,"/analysis/plots/UMAP_HCA_leftventricle_TCF15_DensityPlot.pdf"), width = 6, height = 5)

```


#-) Extended Figure 1e
##8. Dotplot CTRL.LV with TCF15 expression

```{r}
DO.Dotplot(Human_HF_CTRLlv,
    group.by.x = "annotation",
    Feature = "TCF15")

ggsave(paste0(outputFolder, "/Dotplot_TCF15_Human_CTRL.svg"), width = 5, height = 6)
```


#-) Extended Data Figure 2a

##9. HCA Analysis

```{r}
#read data from HCA
cellxgene <- read.csv(
  "/mnt/mariano/scStorage/Mariano/forElvin/2026/request_260123_TCF15_various_tissues/HCA_tissue_data/HCA_CELLxGENE_gene_expression_012326.csv",
  comment.char = "#"
)

head(cellxgene)
unique(cellxgene$Tissue)

cellxgene$pct_expressed <- cellxgene$Number.of.Cells.Expressing.Genes / cellxgene$Cell.Count *100

#check the heart endothelial cell expression
tmp <- cellxgene[cellxgene$Tissue == "heart" & !is.na(cellxgene$Expression),]

# preparing some tissues (data has 53) and cell types (data has 865)

core_tissues <- c(
  "heart",
  "lung",
  "kidney",
  "liver",
  "brain",

  "adipose tissue",
  "musculature"

)

celltype_map <- tibble::tibble(
  pattern = c(
    "cardiomyocyte",
    "fibroblast",
    "endothelial",
    "pericyte",
    "smooth muscle",
    "macrophage|monocyte",
    "t cell|t-cell",
    "b cell|b-cell|plasma",
    "nk cell",
    "neutrophil",
    "dendritic",
    "epithelial",
    "stem|progenitor"
  ),
  broad_type = c(
    "Cardiomyocytes",
    "Fibroblasts",
    "Endothelial cells",
    "Pericytes",
    "Smooth muscle cells",
    "Myeloid cells",
    "T cells",
    "B cells",
    "NK cells",
    "Neutrophils",
    "Dendritic cells",
    "Epithelial cells",
    "Progenitor cells"
  )
)


cellxgene_filt <- cellxgene %>%
  filter(Tissue %in% core_tissues)

tmp <- cellxgene_filt[
  grepl(
    paste(celltype_map$pattern, collapse = "|"),
    cellxgene_filt$Cell.Type,
    ignore.case = TRUE
  ),
]

#Major Map
major_map <- list(
  "Cardiomyocytes" = c("cardiomyocyte"),
  
  "Fibroblasts" = c(
    "fibroblast", "myofibroblast", "stromal", "mesenchymal"
  ),
  
  "Endothelial cells" = c(
    "endothelial"
  ),
  
  "Pericytes & SMCs" = c(
    "pericyte", "smooth muscle"
  ),
  
  "Epithelial cells" = c(
    "epithelial", "goblet", "club", "ciliated", "tuft", "secretory"
  ),
  
  "Myeloid cells" = c(
    "monocyte", "macrophage", "dendritic", "neutrophil", "mast"
  ),
  
  "Lymphoid cells" = c(
    "t cell", "b cell", "plasma", "nk"
  ),
  
  "Progenitor / Stem cells" = c(
    "progenitor", "stem", "precursor"
  ),
  
  "Neuronal / Glial" = c(
    "neuron", "neural", "glial"
  ),
  
  "Other" = c(".*")
)

library(stringr)

assign_major_type <- function(ct, mapping) {
  ct <- tolower(ct)
  for (grp in names(mapping)) {
    pattern <- paste(mapping[[grp]], collapse = "|")
    if (str_detect(ct, pattern)) {
      return(grp)
    }
  }
  return("Other")
}

tmp$MajorCellType <- vapply(
  tmp$Cell.Type,
  assign_major_type,
  mapping = major_map,
  FUN.VALUE = character(1)
)

tmp %>%
  dplyr::select(Cell.Type, MajorCellType) %>%
  distinct() %>%
  arrange(MajorCellType)

plot_df <- tmp %>%
  filter(Gene.Symbol == "TCF15") %>%
  group_by(Tissue, MajorCellType) %>%
  summarise(
    mean_expr = mean(Expression, na.rm = TRUE),
    pct_expr = mean(pct_expressed, na.rm = TRUE),
    .groups = "drop"
  )

plot_df <- plot_df[complete.cases(plot_df),]

dot.size = c(1, 6)
midpoint = 0.5
point_stroke = 0.2
limits_colorscale = NULL
nudge_x = 0.3
nudge_y = 0.2
exp.title <- "Mean expression in group"
fill.values <- plot_df$mean_expr

#filter plot_df by the wanted cell types
plot_df <- plot_df[plot_df$MajorCellType %in% c("Cardiomyocytes", "Endothelial cells", "Fibroblasts", "Pericytes & SMCs"),]


library(ggplot2)
pmain <- ggplot(plot_df, aes(x = MajorCellType, y = Tissue)) +
  theme_bw(base_size = 14) +
  ggplot2::xlab("") + ggplot2::ylab("") + ggplot2::coord_fixed(clip = "off") +
  ggplot2::theme(plot.margin = ggplot2::margin(
    t = 1,
    r = 1,
    b = 1,
    l = 1,
    unit = "cm"
  ),
    axis.text = ggplot2::element_text(color = "black"),
    legend.direction = "horizontal",
    axis.text.x = element_text(
      color = "black",
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 14,
      family = "Helvetica"
    ),
    axis.text.y = element_text(
      color = "black",
      size = 14,
      family = "Helvetica"
    ),
    axis.title.x = element_text(
      color = "black",
      size = 14,
      family = "Helvetica"
    ),
    axis.title = element_text(
      size = 14,
      color = "black",
      family = "Helvetica"
    ),
    plot.title = element_text(
      size = 14,
      hjust = 0.5,
      face = "bold",
      family = "Helvetica"
    ),
    plot.subtitle = element_text(
      size = 14,
      hjust = 0,
      family = "Helvetica"
    ),
    axis.line = element_line(color = "black"),
    strip.text.x = element_text(
      size = 14,
      color = "black",
      family = "Helvetica",
      face = "bold"
    ),
    legend.text = element_text(
      size = 10,
      color = "black",
      family = "Helvetica"
    ),
    legend.title = element_text(
      size = 10,
      color = "black",
      family = "Helvetica",
      hjust = 0
    ),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    
  )
guides.layer <- ggplot2::guides(
  fill = ggplot2::guide_colorbar(
    title = exp.title,
    title.position = "top",
    title.hjust = 0.5,
    barwidth = unit(3.8, "cm"),
    barheight = unit(0.5, "cm"),
    frame.colour = "black",
    frame.linewidth = 0.3,
    ticks.colour = "black",
    order = 2
  ),
  size = ggplot2::guide_legend(
    title = "Fraction of cells \n in group (%)",
    title.position = "top",
    title.hjust = 0.5,
    label.position = "bottom",
    override.aes = list(color = "black", fill = "grey50"),
    keywidth = ggplot2::unit(0.5, "cm"),
    order = 1
  )
)
dot.col <- c("#fff5f0", "#990000")
gradient_colors <- c("#fff5f0", "#fcbba1", "#fc9272", "#fb6a4a", "#990000")

if (length(dot.col) == 2) {
  breaks <- (scales::breaks_extended(n = 5))(range(fill.values))
  if (is.null(limits_colorscale)) {
    limits_colorscale <- c(min(range(fill.values)) * 
                             0.99, max(range(fill.values)) * 1.01)
  }
  if (max(breaks) > max(limits_colorscale)) {
    limits_colorscale[length(limits_colorscale)] <- breaks[length(breaks)]
  }
  pmain <- pmain + ggplot2::scale_fill_gradientn(colours = gradient_colors, 
                                                 breaks = breaks, limits = limits_colorscale)
} else {
  pmain <- pmain + ggplot2::scale_fill_gradient2(low = dot.col[1], 
                                                 mid = dot.col[2], high = dot.col[3], midpoint = midpoint, 
                                                 name = "Gradient")
}


pmain <- pmain +
  geom_point(aes(fill = mean_expr, size = pct_expr),
             shape = 21,
             stroke = point_stroke) +
  guides.layer +
  facet_grid(cols = vars("TCF15")) +
  ggplot2::scale_size(range = c(dot.size[1], dot.size[2]))+
  theme(panel.spacing = unit(0, "lines"))

ggsave("/mnt/mariano/scStorage/Mariano/forElvin/2026/request_260123_TCF15_various_tissues/HCA_tissue_data/Dotplot_HCA_tissues_major_celltypes_sub.svg", width = 5.5, height = 5.5)



```

#-) Extended Data Figure 2b

##10. MCA Analysis

```{r}
#read data from HCA
cellxgene <- read.csv(
  "/mnt/mariano/scStorage/Mariano/forElvin/2026/request_260123_TCF15_various_tissues/MCA_tissue_data/MCA_CELLxGENE_gene_expression_012326.csv",
  comment.char = "#"
)

head(cellxgene)
unique(cellxgene$Tissue)

cellxgene$pct_expressed <- cellxgene$Number.of.Cells.Expressing.Genes / cellxgene$Cell.Count *100

#check the heart endothelial cell expression
tmp <- cellxgene[cellxgene$Tissue == "heart" & !is.na(cellxgene$Expression),]

# preparing some tissues (data has 53) and cell types (data has 865)

core_tissues <- c(
  "heart",
  "lung",
  "kidney",
  "liver",
  "brain",
  "adipose tissue",
  "musculature"

)

celltype_map <- tibble::tibble(
  pattern = c(
    "cardiomyocyte|cardiac muscle cell",
    "fibroblast",
    "endothelial",
    "pericyte",
    "smooth muscle",
    "macrophage|monocyte",
    "t cell|t-cell",
    "b cell|b-cell|plasma",
    "nk cell",
    "neutrophil",
    "dendritic",
    "epithelial",
    "stem|progenitor"
  ),
  broad_type = c(
    "Cardiomyocytes",
    "Fibroblasts",
    "Endothelial cells",
    "Pericytes",
    "Smooth muscle cells",
    "Myeloid cells",
    "T cells",
    "B cells",
    "NK cells",
    "Neutrophils",
    "Dendritic cells",
    "Epithelial cells",
    "Progenitor cells"
  )
)


cellxgene_filt <- cellxgene %>%
  filter(Tissue %in% core_tissues)

tmp <- cellxgene_filt[
  grepl(
    paste(celltype_map$pattern, collapse = "|"),
    cellxgene_filt$Cell.Type,
    ignore.case = TRUE
  ),
]

#Major Map
major_map <- list(
  "Cardiomyocytes" = c("cardiomyocyte", "cardiac muscle cell"),
  
  "Fibroblasts" = c(
    "fibroblast", "myofibroblast", "stromal", "mesenchymal"
  ),
  
  "Endothelial cells" = c(
    "endothelial"
  ),
  
  "Pericytes & SMCs" = c(
    "pericyte", "smooth muscle"
  ),
  
  "Epithelial cells" = c(
    "epithelial", "goblet", "club", "ciliated", "tuft", "secretory"
  ),
  
  "Myeloid cells" = c(
    "monocyte", "macrophage", "dendritic", "neutrophil", "mast"
  ),
  
  "Lymphoid cells" = c(
    "t cell", "b cell", "plasma", "nk"
  ),
  
  "Progenitor / Stem cells" = c(
    "progenitor", "stem", "precursor"
  ),
  
  "Neuronal / Glial" = c(
    "neuron", "neural", "glial"
  ),
  
  "Other" = c(".*")
)

library(stringr)

assign_major_type <- function(ct, mapping) {
  ct <- tolower(ct)
  for (grp in names(mapping)) {
    pattern <- paste(mapping[[grp]], collapse = "|")
    if (str_detect(ct, pattern)) {
      return(grp)
    }
  }
  return("Other")
}

tmp$MajorCellType <- vapply(
  tmp$Cell.Type,
  assign_major_type,
  mapping = major_map,
  FUN.VALUE = character(1)
)

tmp %>%
  select(Cell.Type, MajorCellType) %>%
  distinct() %>%
  arrange(MajorCellType)

plot_df <- tmp %>%
  filter(Gene.Symbol == "Tcf15") %>%
  group_by(Tissue, MajorCellType) %>%
  summarise(
    mean_expr = mean(Expression, na.rm = TRUE),
    pct_expr = mean(pct_expressed, na.rm = TRUE),
    .groups = "drop"
  )

plot_df <- plot_df[complete.cases(plot_df),]

dot.size = c(1, 6)
midpoint = 0.5
point_stroke = 0.2
limits_colorscale = NULL
nudge_x = 0.3
nudge_y = 0.2
exp.title <- "Mean expression in group"
fill.values <- plot_df$mean_expr

#filter plot_df by the wanted cell types
plot_df <- plot_df[plot_df$MajorCellType %in% c("Cardiomyocytes", "Endothelial cells", "Fibroblasts", "Pericytes & SMCs"),]


library(ggplot2)
pmain <- ggplot(plot_df, aes(x = MajorCellType, y = Tissue)) +
  theme_bw(base_size = 14) +
  ggplot2::xlab("") + ggplot2::ylab("") + ggplot2::coord_fixed(clip = "off") +
  ggplot2::theme(plot.margin = ggplot2::margin(
    t = 1,
    r = 1,
    b = 1,
    l = 1,
    unit = "cm"
  ),
    axis.text = ggplot2::element_text(color = "black"),
    legend.direction = "horizontal",
    axis.text.x = element_text(
      color = "black",
      angle = 90,
      hjust = 1,
      vjust = 0.5,
      size = 14,
      family = "Helvetica"
    ),
    axis.text.y = element_text(
      color = "black",
      size = 14,
      family = "Helvetica"
    ),
    axis.title.x = element_text(
      color = "black",
      size = 14,
      family = "Helvetica"
    ),
    axis.title = element_text(
      size = 14,
      color = "black",
      family = "Helvetica"
    ),
    plot.title = element_text(
      size = 14,
      hjust = 0.5,
      face = "bold",
      family = "Helvetica"
    ),
    plot.subtitle = element_text(
      size = 14,
      hjust = 0,
      family = "Helvetica"
    ),
    axis.line = element_line(color = "black"),
    strip.text.x = element_text(
      size = 14,
      color = "black",
      family = "Helvetica",
      face = "bold"
    ),
    legend.text = element_text(
      size = 10,
      color = "black",
      family = "Helvetica"
    ),
    legend.title = element_text(
      size = 10,
      color = "black",
      family = "Helvetica",
      hjust = 0
    ),
    legend.position = "right",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    
  )
guides.layer <- ggplot2::guides(
  fill = ggplot2::guide_colorbar(
    title = exp.title,
    title.position = "top",
    title.hjust = 0.5,
    barwidth = unit(3.8, "cm"),
    barheight = unit(0.5, "cm"),
    frame.colour = "black",
    frame.linewidth = 0.3,
    ticks.colour = "black",
    order = 2
  ),
  size = ggplot2::guide_legend(
    title = "Fraction of cells \n in group (%)",
    title.position = "top",
    title.hjust = 0.5,
    label.position = "bottom",
    override.aes = list(color = "black", fill = "grey50"),
    keywidth = ggplot2::unit(0.5, "cm"),
    order = 1
  )
)
dot.col <- c("#fff5f0", "#990000")
gradient_colors <- c("#fff5f0", "#fcbba1", "#fc9272", "#fb6a4a", "#990000")

if (length(dot.col) == 2) {
  breaks <- (scales::breaks_extended(n = 5))(range(fill.values))
  if (is.null(limits_colorscale)) {
    limits_colorscale <- c(min(range(fill.values)) * 
                             0.99, max(range(fill.values)) * 1.01)
  }
  if (max(breaks) > max(limits_colorscale)) {
    limits_colorscale[length(limits_colorscale)] <- breaks[length(breaks)]
  }
  pmain <- pmain + ggplot2::scale_fill_gradientn(colours = gradient_colors, 
                                                 breaks = breaks, limits = limits_colorscale)
} else {
  pmain <- pmain + ggplot2::scale_fill_gradient2(low = dot.col[1], 
                                                 mid = dot.col[2], high = dot.col[3], midpoint = midpoint, 
                                                 name = "Gradient")
}

pmain <- pmain +
  geom_point(aes(fill = mean_expr, size = pct_expr),
             shape = 21,
             stroke = point_stroke) +
  guides.layer +
  facet_grid(cols = vars("Tcf15")) +
  ggplot2::scale_size(range = c(dot.size[1], dot.size[2]))+
  theme(panel.spacing = unit(0, "lines"))

ggsave("/mnt/mariano/scStorage/Mariano/forElvin/2026/request_260123_TCF15_various_tissues/MCA_tissue_data/Dotplot_MCA_tissues_major_celltypes_sub.svg", width = 5.5, height = 5.5)

```


#-) Extended Data Figure 10b & 10d

##11. Tabula muris cell type summary
```{r}

tm_seu <- readRDS(file = "/mnt/mariano/scStorage/Mariano/forElvin/2026/request_260427_HCA_and_EndMA_marker_analysis/objects/tabula_muris.rds")
# preparing some tissues (data has 53) and cell types (data has 865)
tissues_obj <- unique(tm_seu$tissue)

core_tissues <- unique(tm_seu$tissue) # take all available tissues in the object

# available cell types
ct_objects <- unique(tm_seu$cell_type)

#alternative list
mapping_table <- tibble::tibble(
  cell_type = c(
    # Cardiomyocytes
    "cardiac muscle cell",

    # Fibroblasts
    "fibroblast",
    "fibroblast of cardiac tissue",
    "fibroblast of lung",
    "pulmonary interstitial fibroblast",

    # Endothelial cells
    "endothelial cell",
    "endothelial cell of coronary artery",
    "endothelial cell of hepatic sinusoid",
    "endothelial cell of lymphatic vessel",
    "vein endothelial cell",
    "kidney capillary endothelial cell",
    "fenestrated endothelial cell",
    "endocardial cell",
    "kidney cortex artery cell",

    # Pericytes
    "pericyte",

    # Smooth muscle
    "smooth muscle cell",
    "smooth muscle cell of the pulmonary artery",
    "smooth muscle cell of trachea",
    "bronchial smooth muscle cell",

    # Myeloid cells
    "macrophage",
    "alveolar macrophage",
    "lung macrophage",
    "kupffer cell",
    "monocyte",
    "classical monocyte",
    "non-classical monocyte",
    "intermediate monocyte",
    "myeloid cell",
    "myeloid leukocyte",
    "granulocyte",
    "granulocytopoietic cell",
    "promonocyte",
    "mast cell",
    "basophil",
    "professional antigen presenting cell",
    "macrophage dendritic cell progenitor",

    # T cells
    "t cell",
    "naive t cell",
    "cd4-positive, alpha-beta t cell",
    "cd8-positive, alpha-beta t cell",
    "regulatory t cell",
    "mature nk t cell",
    "immature nk t cell",
    "immature t cell",
    "thymocyte",
    "dn3 thymocyte",
    "dn4 thymocyte",
    "double negative thymocyte",

    # B cells
    "b cell",
    "naive b cell",
    "immature b cell",
    "precursor b cell",
    "late pro-b cell",
    "plasma cell",

    # NK cells
    "natural killer cell",

    # Neutrophils
    "neutrophil",

    # Dendritic cells
    "dendritic cell",
    "myeloid dendritic cell",
    "plasmacytoid dendritic cell",

    # Epithelial cells
    "epithelial cell",
    "keratinocyte",
    "keratinocyte stem cell",
    "basal cell",
    "basal cell of epidermis",
    "epidermal cell",
    "epithelial cell of large intestine",
    "enterocyte of epithelium of large intestine",
    "large intestine goblet cell",
    "intestinal crypt stem cell",
    "duct epithelial cell",
    "pancreatic ductal cell",
    "pancreatic acinar cell",
    "club cell",
    "multiciliated columnar cell of tracheobronchial tree",
    "basal epithelial cell of tracheobronchial tree",
    "pulmonary alveolar type 2 cell",
    "pulmonary neuroendocrine cell",
    "neuroendocrine cell",
    "tuft cell",
    "bladder urothelial cell",
    "bladder cell",
    "luminal epithelial cell of mammary gland",

    # Progenitor / stem
    "hematopoietic stem cell",
    "hematopoietic precursor cell",
    "erythroid progenitor cell, mammalian",
    "megakaryocyte-erythroid progenitor cell",
    "proerythroblast",
    "erythroblast",
    "mesenchymal stem cell",
    "mesenchymal stem cell of adipose tissue",
    "stem cell of epidermis",

    # Extra (still mapped to closest category)
    "stromal cell",
    "mesenchymal cell",
    "adventitial cell",
    "hepatic stellate cell",
    "pancreatic stellate cell",
    "chondrocyte",
    "skeletal muscle satellite cell",
    "cell of skeletal muscle",
    "schwann cell",
    "cardiac neuron",
    "kidney cell",
    "kidney proximal convoluted tubule epithelial cell",
    "kidney distal convoluted tubule epithelial cell",
    "kidney loop of henle ascending limb epithelial cell",
    "kidney loop of henle thick ascending limb epithelial cell",
    "kidney collecting duct principal cell",
    "kidney collecting duct epithelial cell",
    "kidney proximal straight tubule epithelial cell",
    "epithelial cell of proximal tubule",
    "mesangial cell",
    "hepatocyte",
    "type b pancreatic cell",
    "pancreatic a cell",
    "pancreatic d cell",
    "pancreatic pp cell",
    "erythrocyte",
    "blood cell",
    "leukocyte",
    "lymphocyte",
    "langerhans cell",
    "plasmatocyte"
  )
) %>%
  mutate(
    broad_type = case_when(
      cell_type %in% c("cardiac muscle cell") ~ "Cardiomyocytes",

      grepl("fibroblast", cell_type) ~ "Fibroblasts",

      grepl("endothelial|endocardial", cell_type) ~ "Endothelial cells",

      grepl("smooth muscle|pericyte", cell_type) ~ "Pericytes & SMCs",

      grepl("macrophage|monocyte|myeloid|granul|mast|basophil|kupffer|dendritic|neutrophils", cell_type) ~ "Myeloid cells",

      grepl("t cell|thymocyte|b cell|natural killer", cell_type) ~ "Lymphoid cells",

      grepl("epithelial|keratinocyte|urothelial|enterocyte|duct|acinar|club|alveolar|neuroendocrine|tuft|goblet", cell_type) ~ "Epithelial cells",

      grepl("stem|progenitor|precursor|erythroblast", cell_type) ~ "Progenitor cells",

      TRUE ~ NA_character_
    )
)


#assign the new annotation
tm_seu@meta.data <- tm_seu@meta.data %>%
  left_join(mapping_table, by = "cell_type")

tm_seu$broad_type <- as.character(tm_seu$broad_type)
table(tm_seu$broad_type, useNA = "ifany")

#subset by only broad annotated assigned cells
tm_seu_sub <- DO.Subset(tm_seu, ident = "broad_type", ident_name = names(table(tm_seu$broad_type)))

#Cd36, Cav1
DO.Dotplot(
    tm_seu_sub,
    Feature = "Cd36",
    group.by.x = "broad_type",
    group.by.y = "tissue",
    stats_y = FALSE
)



tm_seu_sub_pre <- DO.Subset(tm_seu_sub, ident = "tissue", ident_name = tissue_keep)


table(tm_seu_sub_pre$broad_type, tm_seu_sub_pre$tissue)

DO.Dotplot(
    tm_seu_sub_pre,
    Feature = "Cav1",
    group.by.x = "broad_type",
    group.by.y = "tissue",
    stats_y = FALSE,
    point_stroke = 0.1
)

ggsave(paste0(outputFolder, "/analysis/plots/Tabula_muris_Dotplot_Cav1_various_tissues.pdf"), width = 6.5, height = 5.5)
```

#-) Extended Data Figure 10c & 10e

##12. Recreate UMAP visualisations from Elvin web based
```{r}
# Cav1 & Cd36 in heart
tm_seu_sub_pre_heart <- DO.Subset(tm_seu_sub, ident = "tissue", ident_name = "heart")

tm_seu_sub_pre_heart <- NormalizeData(object = tm_seu_sub_pre_heart)
tm_seu_sub_pre_heart <- ScaleData(object = tm_seu_sub_pre_heart)
tm_seu_sub_pre_heart <- FindNeighbors(object = tm_seu_sub_pre_heart, reduction = "PCA", dims = 1:10)
tm_seu_sub_pre_heart <- FindClusters(tm_seu_sub_pre_heart, resolution = 0.3, algorithm = 4, random.seed = 1234)
tm_seu_sub_pre_heart <- RunUMAP(object = tm_seu_sub_pre_heart, reduction = "PCA", dims = 1:10, spread = 1, min.dist = 0.3, reduction.name = "umap")

p1 <- DO.UMAP(
    tm_seu_sub_pre_heart,
    reduction = "umap",
    group.by = "broad_type",
    legend.position = "none"
)
ggsave(plot = p1, filename = paste0(outputFolder, "/analysis/plots/UMAP_tabula_muris_heart_celltype.pdf"), width = 6, height = 5)
p2 <- DO.UMAP(
    tm_seu_sub_pre_heart,
    features = "Cd36",
    DensityPlot = TRUE,
    reduction = "umap",
    density_quantile_threshold = 0.75,
    legend.position = "right"
)
ggsave(plot = p2, filename = paste0(outputFolder, "/analysis/plots/UMAP_tabula_muris_heart_Cd36.pdf"), width = 6, height = 5)

p3 <- DO.UMAP(
    tm_seu_sub_pre_heart,
    features = "Cav1",
    DensityPlot = TRUE,
    reduction = "umap",
    density_quantile_threshold = 0.75,
    legend.position = "right"
)
ggsave(plot = p3, filename = paste0(outputFolder, "/analysis/plots/UMAP_tabula_muris_heart_Cav1.pdf"), width = 6, height = 5)

# ggpubr::ggarrange(plotlist = list(p1,p2,p3), ncol = 3, nrow = 1)

# Cav1 & Cd36 in limb muscle
tm_seu_sub_pre_lm <- DO.Subset(tm_seu_sub, ident = "tissue", ident_name = "limb muscle")
tm_seu_sub_pre_lm <- NormalizeData(object = tm_seu_sub_pre_lm)
tm_seu_sub_pre_lm <- ScaleData(object = tm_seu_sub_pre_lm)
tm_seu_sub_pre_lm <- FindNeighbors(object = tm_seu_sub_pre_lm, reduction = "PCA", dims = 1:10)
tm_seu_sub_pre_lm <- FindClusters(tm_seu_sub_pre_lm, resolution = 0.3, algorithm = 4, random.seed = 1234)
tm_seu_sub_pre_lm <- RunUMAP(object = tm_seu_sub_pre_lm, reduction = "PCA", dims = 1:10, spread = 1, min.dist = 0.3, reduction.name = "umap")

p1 <- DO.UMAP(
    tm_seu_sub_pre_lm,
    reduction = "umap",
    group.by = "broad_type",
    legend.position = "none"
)
ggsave(plot = p1, filename = paste0(outputFolder, "/analysis/plots/UMAP_tabula_muris_lm_celltype.pdf"), width = 6, height = 5)
p2 <- DO.UMAP(
    tm_seu_sub_pre_lm,
    features = "Cd36",
    DensityPlot = TRUE,
    reduction = "umap",
    density_quantile_threshold = 0.75,
    legend.position = "right"
)
ggsave(plot = p2, filename = paste0(outputFolder, "/analysis/plots/UMAP_tabula_muris_lm_Cd36.pdf"), width = 6, height = 5)

p3 <- DO.UMAP(
    tm_seu_sub_pre_lm,
    features = "Cav1",
    DensityPlot = TRUE,
    reduction = "umap",
    density_quantile_threshold = 0.75,
    legend.position = "right"
)
ggsave(plot = p3, filename = paste0(outputFolder, "/analysis/plots/UMAP_tabula_muris_lm_Cav1.pdf"), width = 6, height = 5)
```